---
title: "A_WildfireSim_DataCleaning"
author: "Sarah Miller"
date: "2025-05-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(xml2)
library(orthogene)
library(tidyverse)
library(dplyr)
library(janitor)
```

# CTD HUMAN DISEASE DATA CLEANING

## GENERATE CTD GENE LISTS:  Read-in NCBI human gene list and CTD disease-specific (more than just human) gene lists. Rename CTD columns and filter Gene_ID to only those found in the NCBI human gene list in order to remove non-human genes.
THESE CTD GENE LISTS WILL HAVE DUPLICATES!!! Mostly due to heterogeneous Disease_name column.
```{r}
# Read-in NCBI human gene list
NCBI <- read_tsv("ncbi_human.tsv")
colnames(NCBI) <- c("NCBI_GeneID","Symbol","Description","Taxonomy_name","Gene_type","Transcripts","Gene_GroupID","Gene_GroupMethod")


# Read-in all CTD gene lists
# Create list of file names (will use as file path for function below)
input <- list.files(path="1_CTD_DiseaseListsV4",
                      pattern=".csv")

# Make list of file names and clean so you can use to rename list below
input_names <- gsub("IS.csv","", input)

# Read-in files
setwd("1_CTD_DiseaseListsV4")
CTD_input_master <- lapply(input, function(x) {
  read_csv(file = x)
})

# Rename list elements with condition
names(CTD_input_master) <- input_names


# Clean CTD gene lists
# Write function
clean_CTD <- function(data){
  colnames(data) <- c("Gene_symbol","Gene_ID","Disease_name","Disease_ID","Direct_evidence","Inference_network","Inference_score","Reference_count")
  data <- data %>%
    filter(Gene_ID %in% NCBI$NCBI_GeneID) %>%
    mutate(Database = "CTD")
  return(data)
}

# Apply functions
CTD_input_master_cleaned <- lapply(CTD_input_master, clean_CTD)
```

# Filter to CTD disease-specific lists
```{r}
# ASTHMA
asthma <- data.frame(CTD_input_master_cleaned[["CTD_asthma"]]) %>%
  mutate(Origin_file = "CTD_asthmaIS.csv",
         Disease = "Asthma")
# COPD
copd <- data.frame(CTD_input_master_cleaned[["CTD_copd"]]) %>%
# Filter for COPD, Chronic bronchitis, and Pulmonary emphysema
  filter(Disease_ID == "MESH:D029424" | Disease_ID == "MESH:D029481" | Disease_ID == "MESH:D011656") %>%
  mutate(Origin_file = "CTD_copdIS.csv",
         Disease = "COPD")

# INTERSTITIAL LUNG DISEASE (ILD)
ild <- data.frame(CTD_input_master_cleaned[["CTD_ild"]]) %>%
  mutate(Origin_file = "CTD_ildIS.csv",
         Disease = "ILD")

# Using CTD_respinfIS.csv as master, filter for individual infections of interest
# PNEUMONIA
pneumonia <- data.frame(CTD_input_master_cleaned[["CTD_respinf"]]) %>%
  filter(Disease_ID == "MESH:D011014") %>%
  mutate(Origin_file = "CTD_RespInfIS.csv",
         Disease = "Pneumonia")
# INFLUENZA
influenza <- data.frame(CTD_input_master_cleaned[["CTD_respinf"]]) %>%
  filter(Disease_ID == "MESH:D007251") %>%
  mutate(Origin_file = "CTD_RespInfIS.csv",
         Disease = "Influenza")
# COVID
covid <- data.frame(CTD_input_master_cleaned[["CTD_respinf"]]) %>%
  filter(Disease_ID == "MESH:D000086382") %>%
  mutate(Origin_file = "CTD_RespInfIS.csv",
         Disease = "COVID-19")
#BRONCHITIS
bronchitis <- data.frame(CTD_input_master_cleaned[["CTD_respinf"]]) %>%
  filter(Disease_ID == "MESH:D001991") %>%
  mutate(Origin_file = "CTD_RespInfIS.csv",
         Disease = "Bronchitis")

# RESPIRATORY HYPERSENSITIVITY
hypersensitivity <- data.frame(CTD_input_master_cleaned[["CTD_resphyp"]]) %>%
# Filter for NOT Asthma
  filter(Disease_ID != "MESH:D001249") %>%
  mutate(Origin_file = "CTD_resphypIS.csv",
         Disease = "Respiratory hypersensitivity")
```

# Create list of cleaned CTD lists
```{r}
CTD_master <- list(asthma, copd, ild, pneumonia, influenza, covid, bronchitis, hypersensitivity)
disease_names <- c("Asthma","COPD","ILD","Pneumonia","Influenza","COVID-19","Bronchitis","Hypersensitivity")
names(CTD_master) <- disease_names
```


# MSigDB HUMAN DISEASE DATA CLEANING

## Extract gene names from MSigDB .xml files. Convert MSigDB .xml files to .csv so you can merge with CTD .csv files
You should only have to do this once, immediately after downloading MSigDB origin files from database. Re-run for each file
```{r}
#Extract gene list from GSEA .xml file
human <- read_xml("3_Top300DiseaseLists/1_MSigDBDiseaseLists/HP_BRONCHITIS.xml")
test <- xml_attrs(xml_child(human, 1))[["MEMBERS_SYMBOLIZED"]]
test <- strsplit(test,",")
human1 <- data.frame(test[[1]]) %>%
  rename('human_gene_name' = 'test..1..')
#write_csv(human1, "3_Top300DiseaseLists/1_MSigDBDiseaseLists/HP_BRONCHITIS.csv")
```

## Read-in MSigDB csv files and clean
```{r}
# Read-in all MSigDB gene lists
# Create list of file names (will use as file path for function below)
input <- list.files(path="1_MSigDB_DiseaseLists",
                      pattern=".csv")

# Make list of file names and clean so you can use to rename list below
input_names <- gsub(".csv","", input)

# Read-in files
setwd("1_MSigDB_DiseaseLists")
MSigDB_input_master <- lapply(input, function(x) {
  read_csv(file = x)
})

# Rename list elements with condition
names(MSigDB_input_master) <- input_names


# Clean MSigDB gene lists
# Write function
clean_MSigDB <- function(data){
  data <- data %>%
    mutate(Database = "MSigDB",
           Inference_score = 1000) %>%
    rename("Gene_symbol" = "human_gene_name")
  return(data)
}

# Apply functions
MSigDB_input_master_cleaned <- lapply(MSigDB_input_master, clean_MSigDB)
```

## Generate disease-specific MSigDB lists
```{r}
# ASTHMA
asthma <- data.frame(MSigDB_input_master_cleaned[["HP_ASTHMA"]]) %>%
  mutate(Origin_file = "HP_ASTHMA.csv")
asthma <- bind_rows(asthma, data.frame(MSigDB_input_master_cleaned[["KEGG_ASTHMA"]] %>%
                                         mutate(Origin_file = "KEGG_ASTHMA.csv")))
asthma <- asthma %>% mutate(Disease = "Asthma")

# COPD
copd <- data.frame(MSigDB_input_master_cleaned[["HP_CHRONIC_PULMONARY_OBSTRUCTION"]]) %>%
  mutate(Origin_file = "HP_CHRONIC_PULMONARY_OBSTRUCTION.csv")
copd <- bind_rows(copd, data.frame(MSigDB_input_master_cleaned[["HP_CHRONIC_BRONCHITIS"]] %>%
                                         mutate(Origin_file = "HP_CHRONIC_BRONCHITIS.csv")))
copd <- bind_rows(copd, data.frame(MSigDB_input_master_cleaned[["HP_EMPHYSEMA"]] %>%
                                         mutate(Origin_file = "HP_EMPHYSEMA.csv")))
copd <- copd %>% mutate(Disease = "COPD")

# ILD
ild <- data.frame(MSigDB_input_master_cleaned[["HP_PULMONARY_FIBROSIS"]]) %>%
  mutate(Origin_file = "HP_PULMONARY_FIBROSIS.csv")
ild <- bind_rows(ild, data.frame(MSigDB_input_master_cleaned[["HP_ABNORMAL_PULMONARY_INTERSTITIAL_MORPHOLOGY"]] %>%
                                         mutate(Origin_file = "HP_ABNORMAL_PULMONARY_INTERSTITIAL_MORPHOLOGY.csv")))
ild <- ild %>% mutate(Disease = "ILD")

# PNEUMONIA
pneumonia <- data.frame(MSigDB_input_master_cleaned[["HP_PNEUMONIA"]]) %>%
  mutate(Origin_file = "HP_PNEUMONIA.csv")
pneumonia <- bind_rows(pneumonia, data.frame(MSigDB_input_master_cleaned[["HP_ACUTE_INFECTIOUS_PNEUMONIA"]] %>%
                                         mutate(Origin_file = "HP_ACUTE_INFECTIOUS_PNEUMONIA.csv")))
pneumonia <- pneumonia %>% mutate(Disease = "Pneumonia")

# INFLUENZA
influenza <- data.frame(MSigDB_input_master_cleaned[["REACTOME_INFLUENZA_INFECTION"]]) %>%
  mutate(Origin_file = "REACTOME_INFLUENZA_INFECTION.csv",
         Disease = "Influenza")

# COVID
covid <- data.frame(MSigDB_input_master_cleaned[["WP_SARS_CORONAVIRUS_AND_INNATE_IMMUNITY"]]) %>%
  mutate(Origin_file = "WP_SARS_CORONAVIRUS_AND_INNATE_IMMUNITY.csv",
         Disease = "COVID-19")

# BRONCHITIS
bronchitis <- data.frame(MSigDB_input_master_cleaned[["HP_BRONCHITIS"]]) %>%
  mutate(Origin_file = "HP_BRONCHITIS.csv",
         Disease = "Bronchitis")

# RESPIRATORY HYPERSENSITIVITY
hypersensitivity <- data.frame(MSigDB_input_master_cleaned[["HP_IMMUNOLOGIC_HYPERSENSITIVITY"]]) %>%
  mutate(Origin_file = "HP_IMMUNOLOGIC_HYPERSENSITIVITY.csv",
         Disease = "Respiratory hypersensitivity")
```

# Create list of cleaned MSigDB lists
```{r}
# Create MSigDB master list
MSigDB_master <- list(asthma, copd, ild, pneumonia, influenza, covid, bronchitis, hypersensitivity)
names(MSigDB_master) <- disease_names
```

# HUMAN DISEASE MASTER LIST GENERATION AND CLEANING

## Combine into master list and generate df to clean
```{r}
# Create master CTD and MSigDB list
Disease_master <- c(CTD_master, MSigDB_master)


# Create dataframe of list elements to save as .csv
# Remove list names so they're not added onto column names
names(Disease_master) <- NULL

# Create df to use as basis for bind_rows in for-loop
Disease_master_df <- as.data.frame(Disease_master[1], row.names = NULL, col.names = names(Disease_master[1]))
iteration = 1

# Loop through Disease_master to bind all exposures to main df
for (i in Disease_master) {
  iteration=iteration+1
  input = Disease_master_df
  
  Disease_master_df <- bind_rows(input, as.data.frame(Disease_master[iteration], row.names = NULL, col.names = names(Disease_master[iteration])))
}

# Save dataframe as raw combined human disease list (contains duplicates)
#write.csv(Disease_master_df, "2_MasterGeneLists/Disease_Human_Master_raw.csv", row.names = FALSE)
```

## Remove duplicate Gene_symbol values, keeping whichever has the higher Inference_score (e.g. prioritize keeping the duplicate associated with MSigDB)
```{r}
# Create function to group Gene_symbol by descending Inference_score and then remove duplicated Gene_symbol, keeping row with highest Inference_score.
remove_dups <- function(data_df) {
  data_df <- data_df %>% 
    group_by(Disease, Gene_symbol) %>%
    arrange(-Inference_score) %>%
    filter(!duplicated(Gene_symbol))
  return(data_df)
}

# Apply function to all disease dataframes in a list
Disease_master_df_nondup <- remove_dups(Disease_master_df)

# Save dataframe as raw combined human disease list (contains duplicates)
#write.csv(Disease_master_df_nondup, "2_MasterGeneLists/Disease_Human_Master_nonduplicate.csv", row.names = FALSE)
```

# MOUSE EXPOSURE DATA CLEANING

## Read-in mouse exposure data

```{r}
# Create list of file names (will use as file path for function below)
expo_deseq <- list.files(path="2_MasterGeneLists/StatResults_RUV_Sarah",
                         pattern="_4h_Lung.csv")

# Make list of file names and clean so you can use to rename list below
expo_deseq_names <- gsub(".csv","", expo_deseq)
expo_deseq_names <- gsub("022821_Stat_Results_RUV_","", expo_deseq_names)
expo_deseq_names <- gsub("_4h_Lung","", expo_deseq_names)

# Read-in all mouse exposure DESEeq2 .csv files
setwd("2_MasterGeneLists/StatResults_RUV_Sarah")
expo_deseq_master <- lapply(expo_deseq, function(x) {
  read.table(file = x,
             sep = ',',
             header = TRUE)
})

# Rename list elements with condition
names(expo_deseq_master) <- expo_deseq_names

# Remove LPS from list and names
expo_deseq_master <- expo_deseq_master[-3]
expo_deseq_names <- expo_deseq_names[-3]
```

There are 10 exposures with 17,356 gene probes (e.g. same gene, different transcripts) per exposure.

## Clean raw exposure data in preparation for ortholog conversion

```{r}
# Write function
add_condition <- function(data, name){
  data <- data %>%
    mutate(Exposure = name) %>% # Add column with exposure condition
    separate("gene", into = c("Mouse_gene", NA), sep = "_") #Rename gene column and remove transcript identifier (need this for ortholog conversion)
}

# Apply function
expo_deseq_master <- mapply(add_condition, expo_deseq_master, names(expo_deseq_master), SIMPLIFY = FALSE)
```

There are 10 exposures still with 17,356 gene probes (e.g. same gene, different transcripts) per exposure.

## Bind lists into df

```{r}
# Remove list names so they're not added onto column names
names(expo_deseq_master) <- NULL

# Create df to use as basis for bind_rows in for-loop
expo_deseq_master_df <- as.data.frame(expo_deseq_master[1], row.names = NULL, col.names = names(expo_deseq_master[1]))
iteration = 1

# Loop through expo_deseq_master to bind all exposures to main df
for (i in expo_deseq_master) {
  iteration=iteration+1
  input = expo_deseq_master_df
  
  expo_deseq_master_df <- bind_rows(input, as.data.frame(expo_deseq_master[iteration], row.names = NULL, col.names = names(expo_deseq_master[iteration])))
}
```

The new data frame contains 173,560 objects because it bound all 17,356 gene probes (e.g. same gene, different transcripts) per exposure.

## Convert mouse exposure genes to human orthologs.

```{r}
# Make df of all unique Mouse_gene in exposure data
expo_genes_MOUSE <- data.frame(
  Mouse_gene = character(),
  stringsAsFactors = FALSE)
expo_genes_MOUSE <- unique(expo_deseq_master_df["Mouse_gene"])

# Make function for converting orthologs 
mouse_to_human <- function(data_df) {
  
  # Prepare input (function needs dummy column to put new genes)
  input_df <- data_df %>%
    mutate(dummy = NA) %>%
    data.frame()

  # Run ortholog conversion
  human_gene_df <- orthogene::convert_orthologs(
    input_df[1:nrow(input_df),],
    gene_input = "Mouse_gene",
    gene_output = "columns",
    standardise_genes = FALSE,
    input_species = "mouse",
    output_species = "human",
    method = "homologene",
    drop_nonorths = TRUE,
    non121_strategy = "drop_both_species",
    mthreshold = Inf,
    as_sparse = FALSE,
    sort_rows = FALSE,
    verbose = TRUE) 
  
  # Clean ortholog conversion output
  human_gene_df <- human_gene_df %>%
    rename("ortholog_gene" = "Human_gene", "input_gene" = "Mouse_gene")

  return(human_gene_df)
  
}

# Apply function to generate human exposure data genes
expo_genes_HUMAN <- mouse_to_human(expo_genes_MOUSE)

# Match human genes back to exposure data and remove rows where no human match was made
expo_deseq_master_df_HUMAN <- left_join(expo_deseq_master_df, expo_genes_HUMAN, by = "Mouse_gene") %>%
  relocate(Human_gene, .after = Mouse_gene)
expo_deseq_master_df_HUMAN <- expo_deseq_master_df_HUMAN[!is.na(expo_deseq_master_df_HUMAN$Human_gene), ]

# Save human version of exposure data (Remember we haven't removed transcripts yet)
#write.csv(expo_deseq_master_df_HUMAN, "2_MasterGeneLists/Exposure_Human_Master_raw.csv", row.names = FALSE)
```


# DESCRIPTIVE STATISTICS AND PLOTS ASSOCIATED WITH CONVERSIONS

## Exposure data: Determine how many genes were dropped in mouse-to-human conversion by exposure

```{r}
# Count mouse genes
count_mouse_genes <- function(df) {
  df %>%
    group_by(Exposure) %>%
    summarise(
      Original_mouse = sum(!is.na(Mouse_gene)),
      Original_mouse_unique = n_distinct(Mouse_gene[!is.na(Mouse_gene)]),
      .groups = "drop"
    )
}
result1 <- count_mouse_genes(expo_deseq_master_df)
print(result1)

# Count human genes
count_human_genes <- function(df) {
  df %>%
    group_by(Exposure) %>%
    summarise(
      Converted_human = sum(!is.na(Human_gene)),
      Converted_human_unique = n_distinct(Human_gene[!is.na(Human_gene)]),
      .groups = "drop"
    )
}
result2 <- count_human_genes(expo_deseq_master_df_HUMAN)
print(result2)

# Take difference to find number dropped
dropped <- full_join(result1, result2, by = "Exposure") %>%
  mutate(Dropped_genes_total = (Original_mouse - Converted_human),
         Perc_dropped_total = (Dropped_genes_total / Original_mouse) * 100,
         Perc_dropped_total = round(Perc_dropped_total, 1),
         Dropped_genes_unique = (Original_mouse_unique - Converted_human_unique),
         Perc_dropped_unique = (Dropped_genes_unique / Original_mouse) * 100,
         Perc_dropped_unique = round(Perc_dropped_unique, 1))
print(dropped)
```

## Disease data: Determine how many duplicate genes were dropped

```{r}
# Count total genes before removing duplicates
count_total_genes <- function(df) {
  df %>%
    group_by(Disease, Database) %>%
    summarise(Total_genes = sum(!is.na(Gene_symbol)), .groups = "drop")
}
result1 <- count_total_genes(Disease_master_df)
print(result1)

# Count non-duplicate genes
count_nondup_genes <- function(df) {
  df %>%
    group_by(Disease, Database) %>%
    summarise(
      Nonduplicate_genes = sum(!is.na(Gene_symbol)), .groups = "drop")
}
result2 <- count_nondup_genes(Disease_master_df_nondup)
print(result2)

# Take difference to find number dropped
dropped <- full_join(result1, result2, by = c("Disease", "Database")) %>%
  mutate(Dropped_genes = (Total_genes - Nonduplicate_genes),
         Perc_dropped = (Dropped_genes / Total_genes) * 100,
         Perc_dropped = round(Perc_dropped, 1))
print(dropped)
```

## Disease data: Generate UpSet plots to look at how many genes are shared across groups

```{r}
# Clean data for comparisons
# Select for disease, filter for genes, rename column to disease (will need to be the same for intersecting and pulling shared gene names later), and change gene names to lowercase
WildfireSim_asthma <- Disease_master_df_nondup %>%
  filter(Disease == "Asthma") %>%
  ungroup() %>%
  select(Gene_symbol)
WildfireSim_bronch <- Disease_master_df_nondup %>%
  filter(Disease == "Bronchitis") %>%
  ungroup() %>%
  select(Gene_symbol)
WildfireSim_copd <- Disease_master_df_nondup %>%
  filter(Disease == "COPD") %>%
  ungroup() %>%
  select(Gene_symbol)
WildfireSim_covid <- Disease_master_df_nondup %>%
  filter(Disease == "COVID-19") %>%
  ungroup() %>%
  select(Gene_symbol)
WildfireSim_ild <- Disease_master_df_nondup %>%
  filter(Disease == "ILD") %>%
  ungroup() %>%
  select(Gene_symbol)
WildfireSim_influ <- Disease_master_df_nondup %>%
  filter(Disease == "Influenza") %>%
  ungroup() %>%
  select(Gene_symbol)
WildfireSim_pneum <- Disease_master_df_nondup %>%
  filter(Disease == "Pneumonia") %>%
  ungroup() %>%
  select(Gene_symbol)
WildfireSim_resphyp <- Disease_master_df_nondup %>%
  filter(Disease == "Respiratory hypersensitivity") %>%
  ungroup() %>%
  select(Gene_symbol)
```

## Generate UpSet plots
```{r}
library(ComplexHeatmap)

lt <- list(Asthma = WildfireSim_asthma$Gene_symbol,
           Bronchitis = WildfireSim_bronch$Gene_symbol,
           COPD = WildfireSim_copd$Gene_symbol,
           "COVID-19" = WildfireSim_covid$Gene_symbol,
           ILD = WildfireSim_ild$Gene_symbol,
           Influenza = WildfireSim_influ$Gene_symbol,
           Pneumonia = WildfireSim_pneum$Gene_symbol,
           Hypersensitivity = WildfireSim_resphyp$Gene_symbol)
list_to_matrix(lt)

m1 = make_comb_mat(lt) # the default mode is `distinct`
m2 = make_comb_mat(lt, mode = "intersect")
m3 = make_comb_mat(lt, mode = "union")

# Distinct
p1 <- UpSet(m1,
            set_order = c("Asthma", "Bronchitis", "COPD", "COVID-19", "ILD", "Influenza", "Pneumonia", "Hypersensitivity"), comb_order = order(comb_size(m1)),
            pt_size = unit(5, "mm"), lwd = 3,
            #comb_col = c("#1E88E5", "#D81B60", "#FFC107")[comb_degree(m1)],
            comb_col = c("#CCFFCC", "#99FF99", "#66FF66", "#33CC33", "#009900", "#006600", "#004400", "#002200")[comb_degree(m1)],
            right_annotation = upset_right_annotation(m1,
                                                      gp = gpar(fill = "black")))
p1
p1L <- UpSet(m1,
            set_order = c("Asthma", "Bronchitis", "COPD", "COVID-19", "ILD", "Influenza", "Pneumonia", "Hypersensitivity"), comb_order = order(comb_size(m1)),
            pt_size = unit(5, "mm"), lwd = 3,
            #comb_col = c("#1E88E5", "#D81B60", "#FFC107")[comb_degree(m1)],
            comb_col = c("#CCFFCC", "#99FF99", "#66FF66", "#33CC33", "#009900", "#006600", "#004400", "#002200")[comb_degree(m1)],
            top_annotation = upset_top_annotation(m1, add_numbers = TRUE),
            right_annotation = upset_right_annotation(m1,
                                                      gp = gpar(fill = "black"),
                                                      width = unit(3, "cm"),
                                                      add_numbers = TRUE))
p1L

# Intersect
p2 <- UpSet(m2,
            set_order = c("Asthma", "Bronchitis", "COPD", "COVID-19", "ILD", "Influenza", "Pneumonia", "Hypersensitivity"), comb_order = order(comb_size(m2)),
            pt_size = unit(5, "mm"), lwd = 3,
            comb_col = c("#E6CCFF", "#D1B3FF", "#B266FF", "#9933FF", "#8000FF", "#6600CC", "#4B0082", "#2E003E")[comb_degree(m2)],
            right_annotation = upset_right_annotation(m2,
                                                      gp = gpar(fill = "black")))
p2
p2L <- UpSet(m2,
            set_order = c("Asthma", "Bronchitis", "COPD", "COVID-19", "ILD", "Influenza", "Pneumonia", "Hypersensitivity"), comb_order = order(comb_size(m2)),
            pt_size = unit(5, "mm"), lwd = 3,
            comb_col = c("#E6CCFF", "#D1B3FF", "#B266FF", "#9933FF", "#8000FF", "#6600CC", "#4B0082", "#2E003E")[comb_degree(m2)],
            top_annotation = upset_top_annotation(m2,
                                                  height = unit(4, "cm"),
                                                  add_numbers = TRUE),
            right_annotation = upset_right_annotation(m2,
                                                      gp = gpar(fill = "black"),
                                                      width = unit(3, "cm"),
                                                      add_numbers = TRUE))
p2L
```

## Export UpSet plots
NOTE: Export this PDF to TIFF and upload to BioRender for post-processing.
```{r eval = FALSE}
pdf("3_DescriptiveFigures/250507_Disease_UpSet_Distinct.pdf", height = 6, width = 16)
p1
invisible(dev.off())
pdf("3_DescriptiveFigures/250507_Disease_UpSet_Distinct_labeled.pdf", height = 6, width = 16)
p1L
invisible(dev.off())

pdf("3_DescriptiveFigures/250507_Disease_UpSet_Intersect.pdf", height = 8, width = 20)
p2
invisible(dev.off())
pdf("3_DescriptiveFigures/250507_Disease_UpSet_Intersect_labeled.pdf", height = 8, width = 20)
p2L
invisible(dev.off())
```