---
title: "B_WildfireSim_CountMatrix_MultipleMetrics"
author: "Sarah Miller"
date: "2025-05-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(xml2)
library(tidyverse)
library(dplyr)
library(janitor)
library(ggplot2)
library(gtools)
library(vegan)
#install.packages("GeneSelectR")
library(GeneSelectR)
#Warning: The following Bioconductor packages are required for full functionality of GeneSelectR but are not installed: clusterProfiler, GO.db, simplifyEnrichment. Please install them using BiocManager::install().
#install.packages("simplifyEnrichment")
#BiocManager::install("simplifyEnrichment")
#library(simplifyEnrichment)

GeneNum = 750
date = 260709

# Assign here so can be saved as metadata below
dis_filter <- paste0("Top ", GeneNum, " according to descending Inference_score, where MSigDB genes are assigned Inference_score=1000 and non-duplicate CTD genes are amended up to target ", GeneNum, " genes")
expo_filter <- paste0("lfcSE < quantile(lfcSE, 0.95), top ", GeneNum, " according to descending absolute value of fold-change")
#expo_filter <- paste0("padj < 0.10, top ", GeneNum, " according to descending absolute value of fold-change")

#250508 - Exposure filter: filter(padj < 0.10) / Metric: Jaccard only
#250515 - Exposure filter: filter(lfcSE < quantile(lfcSE, 0.95)) / Metric: Jaccard only
#250519 - Exposure filter: filter(lfcSE < quantile(lfcSE, 0.95)) / Metric: Jaccard, Overlap, Soerensen
#250520 - Exposure filter: filter(padj < 0.10) / Metric: Jaccard, Overlap, Soerensen
#250523 - Exposure filter: filter(lfcSE < quantile(lfcSE, 0.95)) / Metric: Jaccard, Overlap, Soerensen (built by Jessie)
#250709 - Exposure filter: filter(lfcSE < quantile(lfcSE, 0.95)) / Metric: Jaccard, Overlap, Soerensen (built by Jessie) - INCLUDE LPS
```

# GENERATE SIGNATURES AND COUNT MATRIX

## Read-in data frames
```{r}
Disease_raw <- read_csv("2_MasterGeneLists/Disease_Human_Master_nonduplicate.csv")
Exposure_raw <- read_csv("2_MasterGeneLists/Exposure_Human_Master_raw_plusLPS.csv") #NOTE THIS IS EXPOSURE LIST THAT CONTAINS LPS GENES
```

## Clean exposure data in preparation for signature defining and split into exposure-specific list
```{r}
# Function to apply the filter
clean_exposure <- function(data) {
  data <- data %>%
    mutate(FC = logratio2foldchange(log2FoldChange, base = 2)) %>% #Convert log2FC into FC for interpretability
    relocate(last_col(), .after = log2FoldChange) %>% #Arrange FC closer to log2FoldChange because related
    filter(lfcSE < quantile(lfcSE, 0.95)) %>% #Filter lfcSE below 95th percentile
    #select(-Mouse_gene, -baseMean, -lfcSE, -stat, -log2FoldChange, -pvalue) %>% #Remove initial columns we don't need
    dplyr::select(-Mouse_gene, -baseMean, -stat) %>% #Remove initial columns we don't need
    #filter(padj<0.10) %>% #Filter for padj
    group_by(Exposure) %>%
    arrange(Human_gene, desc(abs(FC))) %>% #Order Human_gene such that transcript with largest abs(FC) is listed first
    filter(!duplicated(Human_gene)) #Remove Human_gene transcripts with smaller magnitude abs(FC)
}

# Apply function
input_expo <- clean_exposure(Exposure_raw)

# Split df into exposure-specific list
input_expo_list <- split(input_expo, input_expo$Exposure)
```

## Define exposure signatures by pulling top GeneNum (see top of code) genes by abs(FC) per exposure
```{r}
# Function to apply the filter
define_expo_signatures <- function(data, name) {
  data <- data %>%
    ungroup() %>%
    arrange(desc(abs(FC))) %>% #Order genes by abs(FC) such that top_n() will work
    top_n(as.integer(paste0(GeneNum)), abs(FC)) %>% #Pull top GeneNum genes according to largest abs(FC) magnitude
    #select(-padj, -FC, -Exposure) %>% #Remove columns we don't need
    mutate(!!name := 1) #Add column named by exposure for count matrix
}

# Apply function
signatures_expo <- mapply(define_expo_signatures, input_expo_list, names(input_expo_list), SIMPLIFY = FALSE)
```

## Write out signatures_expo for supplementary data
```{r}
# Clean signatures_expo for writing out as supplementary data
expo_supptable_list <- signatures_expo # Create dummy list so signatures_expo isn't affected for downstream manipulations
names(expo_supptable_list) <- NULL # Remove list names so they're not added onto column names
expo_supptable_list <- expo_supptable_list %>%
  map(~ .x %>% dplyr::select(-8)) # Remove 8th column named for Exposure filled with "1"

# Create df to use as basis for bind_rows in for-loop
expo_supptable <- as.data.frame(expo_supptable_list[1], row.names = NULL, col.names = names(expo_supptable_list[1]))
iteration = 1

# Loop through expo_supptable_list to bind all exposures to main df
for (i in expo_supptable_list) {
  tryCatch({
    iteration=iteration+1
    input = expo_supptable
  
    #expo_supptable <- full_join(input, as.data.frame(expo_supptable_list[iteration], row.names = NULL), by = join_by(Human_gene))
    expo_supptable <- bind_rows(input, as.data.frame(expo_supptable_list[iteration]))
  }, error=function(e){cat("Error :",conditionMessage(e), "\n")})
}

# Write out expo_supptable
write_csv(expo_supptable, file = paste0("4_CountMatrix_Jaccard/", date, "_", GeneNum, "_plusLPS_CountMatrix_MultiMeasures_SUMMARYSTATS.csv"))
```

## Define exposure signatures by pulling top GeneNum (see top of code) genes by abs(FC) per exposure
```{r}
# Function to apply the filter
clean_define_expo_signatures <- function(data, name) {
  data <- data %>%
    dplyr::select(-log2FoldChange, -FC, -lfcSE, -pvalue,-padj, -Exposure) #Remove columns we don't need
}

# Apply function
signatures_expo <- mapply(clean_define_expo_signatures, signatures_expo, names(signatures_expo), SIMPLIFY = FALSE)
```

## Clean disease data and define signatures by pulling top GeneNum (see top of code) genes by Inference_score
```{r}
# Function to apply the filter
clean_define_dis_signatures <- function(data) {
  data <- data %>%
    dplyr::select(Gene_symbol, Database, Inference_score, Disease) %>%
    dplyr::rename("Human_gene" = "Gene_symbol") %>%
    group_by(Disease) %>%
    top_n(as.integer(paste0(GeneNum)), Inference_score)
}

# Apply function
input_dis <- clean_define_dis_signatures(Disease_raw)

# Split df into disease-specific list
input_dis_list <- split(input_dis, input_dis$Disease)
```

## Add disease-specific count column so disease signatures can be added to count matrix
```{r}
# Function to apply the filter
addcol_dis_signatures <- function(data, name) {
  data <- data %>%
    ungroup() %>%
    dplyr::select(-Database, -Inference_score, -Disease) %>% #Remove columns we don't need
    mutate(!!name := 1) #Add column named by disease for count matrix
}

# Apply function
signatures_dis <- mapply(addcol_dis_signatures, input_dis_list, names(input_dis_list), SIMPLIFY = FALSE)
```

## Create count matrix for saving 
```{r}
# EXPOSURE
# Remove list names so they're not added onto column names
names(signatures_expo) <- NULL

# Create df to use as basis for bind_rows in for-loop
count_matrix <- as.data.frame(signatures_expo[1], row.names = NULL, col.names = names(signatures_expo[1]))
iteration = 1

# Loop through signatures_expo to bind all exposures to main df
for (i in signatures_expo) {
  tryCatch({
    iteration=iteration+1
    input = count_matrix
  
    count_matrix <- full_join(input, as.data.frame(signatures_expo[iteration], row.names = NULL), by = join_by(Human_gene))
  }, error=function(e){cat("Error :",conditionMessage(e), "\n")})
}
# This is going to throw an error message, it's okay (it worked, there should be 2134 genes in count_matrix)

# DISEASE
# Remove list names so they're not added onto column names
names(signatures_dis) <- NULL

iteration = 0
# Loop through signatures_expo to bind all exposures to main df
for (i in signatures_dis) {
  iteration=iteration+1
  input = count_matrix
  
  count_matrix <- full_join(input, as.data.frame(signatures_dis[iteration], row.names = NULL), by = join_by(Human_gene))
}
# This is going to throw an error message, it's okay
```

## Clean count matrix, generate summary statistics, and save
```{r}
# Replace NAs with 0
count_matrix[is.na(count_matrix)] <- 0

# Rename columns and relocate LPS at end of exposures
cols <- c("Human_gene","Flaming_eucalyptus","Smoldering_eucalyptus","LPS","Flaming_peat","Smoldering_peat","Flaming_pine_needles","Smoldering_pine_needles","Flaming_pine","Smoldering_pine","Flaming_red_oak","Smoldering_red_oak","Asthma","Bronchitis","COPD","COVID-19","ILD","Influenza","Pneumonia","Hypersensitivity")
colnames(count_matrix) <- cols
count_matrix <- count_matrix %>%
  relocate(LPS, .before = Asthma)

# Sum across columns to identify shared genes, and sort by highest value
stats_matrix <- as.data.frame(colSums(count_matrix[,2:ncol(count_matrix)]))
colnames(stats_matrix) <- "Genes per signature"
stats_matrix <- rownames_to_column(stats_matrix) %>%
  dplyr::rename("Condition" = "rowname")

# Create df of exposure filter terms to save as metadata
filter_terms <- data.frame(Condition = c("Exposure", "Disease"),
                           Filter = c(paste0(expo_filter), paste0(dis_filter)))

# Create Excel workbook
OUT <- createWorkbook()
addWorksheet(OUT, "count_matrix")
writeData(OUT, sheet = "count_matrix", x = count_matrix, rowNames = FALSE)
addWorksheet(OUT, "signature_stats")
writeData(OUT, sheet = "signature_stats", x = stats_matrix)
addWorksheet(OUT, "filter_terms")
writeData(OUT, sheet = "filter_terms", x = filter_terms)
# Export file
saveWorkbook(OUT, paste0("4_CountMatrix_Jaccard/", date, "_", GeneNum, "_plusLPS_CountMatrix_MultiMeasures.xlsx"), overwrite = TRUE)
```

# CALCULATE SIMILARITY COEFFICIENTS

## Re-define signatures_expo and signatures_dis as lists of gene names so they're formatted correctly to calculate different similarity metrics using calculate_overlap_coefficients()
```{r}
# EXPOSURE
prep_expo_signatures <- function(data) {
  data <- data %>%
    ungroup() %>%
    arrange(desc(abs(FC))) %>% #Order genes by abs(FC) such that top_n() will work
    top_n(as.integer(paste0(GeneNum)), abs(FC)) %>% #Pull top GeneNum genes according to largest abs(FC) magnitude
    dplyr::select(-padj, -FC, -Exposure) #Remove columns we don't need
}

# Apply function
signatures_expo <- mapply(prep_expo_signatures, input_expo_list, SIMPLIFY = FALSE)


# DISEASE
prep_dis_signatures <- function(data) {
  data <- data %>%
    ungroup() %>%
    dplyr::select(-Database, -Inference_score, -Disease) #Remove columns we don't need
}

# Apply function
signatures_dis <- mapply(prep_dis_signatures, input_dis_list, SIMPLIFY = FALSE)
```

## Merge into master list and run multiple similarity coefficients
```{r}
# Make master list and rename using cols assigned above
signatures_total <- c(signatures_expo, signatures_dis)
names(signatures_total) <- cols[2:20] #Changed to 20 from 19 to account for LPS

# Define the three similarity functions
jaccard_similarity <- function(set1, set2) {
  intersect_len <- length(intersect(set1, set2))
  union_len <- length(union(set1, set2))
  return(intersect_len / union_len)
}

dice_coefficient <- function(set1, set2) {
  intersect_len <- length(intersect(set1, set2))
  return(2 * intersect_len / (length(set1) + length(set2)))
}

overlap_coefficient <- function(set1, set2) {
  intersect_len <- length(intersect(set1, set2))
  return(intersect_len / min(length(set1), length(set2)))
}

# Extract sets of genes
gene_sets <- lapply(signatures_total, function(df) unique(df$Human_gene))

# Number of signatures
n <- length(gene_sets)
labels <- names(signatures_total)

# Initialize matrices
jaccard_matrix <- matrix(0, n, n, dimnames = list(labels, labels))
dice_matrix <- matrix(0, n, n, dimnames = list(labels, labels))
overlap_matrix <- matrix(0, n, n, dimnames = list(labels, labels))

# Compute pairwise similarities
for (i in seq_len(n)) {
  for (j in seq_len(n)) {
    set1 <- gene_sets[[i]]
    set2 <- gene_sets[[j]]
    
    jaccard_matrix[i, j] <- jaccard_similarity(set1, set2)
    dice_matrix[i, j] <- dice_coefficient(set1, set2)
    overlap_matrix[i, j] <- overlap_coefficient(set1, set2)
  }
}

# Convert to data frames 
jaccard_df <- as.data.frame(jaccard_matrix) %>%
  rownames_to_column("Condition")
dice_df <- as.data.frame(dice_matrix) %>%
  rownames_to_column("Condition")
overlap_df <- as.data.frame(overlap_matrix) %>%
  rownames_to_column("Condition")
```

## Create and save Excel of all data
```{r}
# Add Excel sheets and write-out similarity matrices
addWorksheet(OUT, "jaccard_matrix")
writeData(OUT, sheet = "jaccard_matrix", x = jaccard_df)
addWorksheet(OUT, "overlap_matrix")
writeData(OUT, sheet = "overlap_matrix", x = overlap_df)
addWorksheet(OUT, "dice_matrix")
writeData(OUT, sheet = "dice_matrix", x = dice_df)
# Export file
saveWorkbook(OUT, paste0("4_CountMatrix_Jaccard/", date, "_", GeneNum, "_plusLPS_CountMatrix_MultiMeasures.xlsx"), overwrite = TRUE)
```

# GENERATE BALLOON PLOTS

# JACCARD

## First, scale jaccard_df within EXPOSURES
```{r}
# Create df to use as basis for bind_rows in for-loop
Index_raw <- jaccard_df[c(1:11),c(1,13)] #To account for LPS, changed columns to 11 from 10 (Exposures) and rows to 13 from 12 so Asthma is pulled
Index_raw$Disease <- colnames(Index_raw)[2]
Index_raw <- Index_raw %>%
  dplyr::rename("Exposure" = "Condition",
                "Index_raw" = "Asthma") %>%
  relocate(Disease, .after = Exposure)

# Define column names you want looped through (note that Asthma is gone because we defined it in the df above)
dis_cols <- c("Bronchitis","COPD","COVID-19","ILD","Influenza","Pneumonia","Hypersensitivity")
iteration = 13 #Changed to 13 from 12 to account for LPS
# Loop through signatures_expo to bind all exposures to main df
for (i in dis_cols) {
  iteration=iteration+1
  input = Index_raw
  
  next_dis <- jaccard_df[c(1:11),c(1,iteration)] #Changed to 11 from 10 to account for LPS
  next_dis$Disease <- colnames(next_dis)[2]
  next_dis <- next_dis %>%
    dplyr::rename("Exposure" = "Condition",
                  "Index_raw" = paste0(i)) %>%
    relocate(Disease, .after = Exposure)
  
  Index_raw <- full_join(input, next_dis)
}

# Create function to min-max scale within columns
minmax <- function(x, na.rm = TRUE) {
    return((x-min(x))/(max(x)-min(x)))
}

# Pivot input_data wider where Exposure conditions become column names
Index_scale <- pivot_wider(Index_raw, names_from = Exposure, values_from = Index_raw)

# Apply function across Exposure columns, pivot back to long, and join back to input_data
Index_scale_minmax <- Index_scale %>%
  mutate(across("Flaming_eucalyptus":"Smoldering_red_oak", minmax)) %>%
  pivot_longer(!Disease, names_to = "Exposure", values_to = "Index_ExpSc")
Index_input <- full_join(Index_raw, Index_scale_minmax)
```

## Create input dataframe for balloon plot
```{r}
# Assign numeric values to each of your groups to assign the position (x-y coordinates) they will appear on the balloon plot. 
row_df <- data.frame(Disease = c("Asthma","COPD","Hypersensitivity","ILD","Bronchitis","COVID-19","Influenza","Pneumonia"), row = c(8,7,6,5,4,3,2,1))
col_df <- data.frame(Exposure = c("Flaming_eucalyptus","Smoldering_eucalyptus","Flaming_peat","Smoldering_peat","Flaming_pine","Smoldering_pine","Flaming_pine_needles","Smoldering_pine_needles","Flaming_red_oak","Smoldering_red_oak","LPS"), col = c(1,2,3,4,5,6,7,8,9,10,11))

# Join row and col positions to the data frame 
input_data <- Index_input %>%
  left_join(row_df, by = "Disease") %>%
  left_join(col_df, by = "Exposure")

# Manually create axes labels (if you want the labels to be different - I think if you don't do this, it will just keep current names)
# Note that these should be in numeric order for the positions on your axis (row position 1 should have its label in the first slot in the vector)
vars_x_axis <- c("Euc_Fl", "Euc_Sm","Peat_Fl","Peat_Sm","Pine_Fl","Pine_Sm","PineNeedles_Fl","PineNeedles_Sm","RedOak_Fl","RedOak_Sm","LPS")
vars_y_axis <- c("Pneumonia","Influenza","COVID-19","Bronchitis","ILD","Hypersensitivity","COPD","Asthma")
```

## Generate balloon plot
Remember to change ggplot(size=) and geom_text(label=)
```{r}
balloonplot <- ggplot(input_data, aes(x=factor(col), y=factor(row), color=Exposure, size=Index_ExpSc)) +
  scale_fill_brewer(palette="Dark2") +
  geom_point() +    # plot as points
  geom_text(aes(label=formatC(as.numeric(as.character(round(Index_ExpSc, 3))), digits = 3, format = "f"), x=col + 0.25), size=4) +   # display the value next to the "balloons"
  scale_size(range = c(1, 5.5)) + # controls range of dot sizes
  scale_x_discrete(breaks=1:length(vars_x_axis), labels=vars_x_axis, position='top') +   
  scale_y_discrete(breaks=1:length(vars_y_axis), labels=vars_y_axis) +                 
  theme_bw() +
  theme(axis.line = element_blank(),            
        axis.title = element_blank(),          
        panel.border = element_blank(),         
        panel.grid.major.x = element_blank(),   
        panel.grid.minor.x = element_blank(),
        legend.position = "none",
        axis.text = element_text(size = 10, color = "black"),
        axis.ticks = element_blank())
balloonplot
```

## Save balloon plot
Note that the dimensions of the figure you write out will influence how close the rows and columns are to each other. These output dimensions worked well for me for my balloon plot with 4 columns and 5 rows.
NOTE: Export this PDF to TIFF and upload to BioRender for post-processing.
```{r}
pdf(paste0("5_Figures/", date, "_", GeneNum, "_plusLPS_BalloonPlot_Jaccard_ExpScaled.pdf"), height = 4, width = 14)
balloonplot
invisible(dev.off())
```

## Second, scale JSIs within DISEASES
```{r}
# Create function to min-max scale within columns
minmax <- function(x, na.rm = TRUE) {
    return((x-min(x))/(max(x)-min(x)))
}

# Pivot input_data wider where Disease conditions become column names
Index_scale <- pivot_wider(Index_raw, names_from = Disease, values_from = Index_raw)

# Apply function across Disease columns, pivot back to long, and join back to input_data
Index_scale_minmax <- Index_scale %>%
  mutate(across("Asthma":"Hypersensitivity", minmax)) %>% #Remember you made this minmax function above
  pivot_longer(!Exposure, names_to = "Disease", values_to = "Index_DisSc")
Index_input <- full_join(Index_input, Index_scale_minmax) %>%
  mutate(Coefficient = "Jaccard")

# Save as Jaccard-specific object
Jaccard_input <- Index_input
```

## Create input dataframe for balloon plot
```{r}
# Assign numeric values to each of your groups to assign the position (x-y coordinates) they will appear on the balloon plot. 
row_df <- data.frame(Exposure = c("Flaming_eucalyptus","Smoldering_eucalyptus","Flaming_peat","Smoldering_peat","Flaming_pine","Smoldering_pine","Flaming_pine_needles","Smoldering_pine_needles","Flaming_red_oak","Smoldering_red_oak","LPS"), row = c(11,10,9,8,7,6,5,4,3,2,1))
col_df <- data.frame(Disease = c("Asthma","COPD","Hypersensitivity","ILD","Bronchitis","COVID-19","Influenza","Pneumonia"), col = c(1,2,3,4,5,6,7,8))

# Join row and col positions to the data frame 
input_data <- Index_input %>%
  left_join(row_df, by = "Exposure") %>%
  left_join(col_df, by = "Disease")

# Manually create axes labels (if you want the labels to be different - I think if you don't do this, it will just keep current names)
# Note that these should be in numeric order for the positions on your axis (row position 1 should have its label in the first slot in the vector)
vars_x_axis <- c("Asthma","COPD","Hypersensitivity","ILD","Bronchitis","COVID-19","Influenza","Pneumonia")
vars_y_axis <- c("LPS","RedOak_Sm","RedOak_Fl","PineNeedles_Sm","PineNeedles_Fl","Pine_Sm","Pine_Fl","Peat_Sm","Peat_Fl","Euc_Sm","Euc_Fl")
```

## Generate balloon plot
Remember to change ggplot(size=) and geom_text(label=)
```{r}
balloonplot <- ggplot(input_data, aes(x=factor(col), y=factor(row), color=Disease, size=Index_DisSc)) +
  scale_fill_brewer(palette="Dark2") +
  geom_point() +    # plot as points
  geom_text(aes(label=formatC(as.numeric(as.character(round(Index_DisSc, 3))), digits = 3, format = "f"), x=col + 0.25), size=4) +   # display the value next to the "balloons"
  scale_size(range = c(1, 5.5)) + # controls range of dot sizes
  scale_x_discrete(breaks=1:length(vars_x_axis), labels=vars_x_axis, position='top') +   
  scale_y_discrete(breaks=1:length(vars_y_axis), labels=vars_y_axis) +                 
  theme_bw() +
  theme(axis.line = element_blank(),            
        axis.title = element_blank(),          
        panel.border = element_blank(),         
        panel.grid.major.x = element_blank(),   
        panel.grid.minor.x = element_blank(),
        legend.position = "none",
        axis.text = element_text(size = 10, color = "black"),
        axis.ticks = element_blank())
balloonplot
```

# Save balloon plot
Note that the dimensions of the figure you write out will influence how close the rows and columns are to each other. These output dimensions worked well for me for my balloon plot with 4 columns and 5 rows.
NOTE: Export this PDF to TIFF and upload to BioRender for post-processing.
```{r}
pdf(paste0("5_Figures/", date, "_", GeneNum, "_plusLPS_BalloonPlot_Jaccard_DisScaled.pdf"), height = 6, width = 11)
balloonplot
invisible(dev.off())
```

# OVERLAP

## First, scale overlap_df within EXPOSURES
```{r}
# Create df to use as basis for bind_rows in for-loop
Index_raw <- overlap_df[c(1:11),c(1,13)] #To account for LPS, changed columns to 11 from 10 (Exposures) and rows to 13 from 12 so Asthma is pulled
Index_raw$Disease <- colnames(Index_raw)[2]
Index_raw <- Index_raw %>%
  dplyr::rename("Exposure" = "Condition",
                "Index_raw" = "Asthma") %>%
  relocate(Disease, .after = Exposure)

# Define column names you want looped through (note that Asthma is gone because we defined it in the df above)
dis_cols <- c("Bronchitis","COPD","COVID-19","ILD","Influenza","Pneumonia","Hypersensitivity")
iteration = 13
# Loop through signatures_expo to bind all exposures to main df
for (i in dis_cols) {
  iteration=iteration+1
  input = Index_raw
  
  next_dis <- overlap_df[c(1:11),c(1,iteration)] #Changed to 11 from 10 to account for LPS
  next_dis$Disease <- colnames(next_dis)[2]
  next_dis <- next_dis %>%
    dplyr::rename("Exposure" = "Condition",
                  "Index_raw" = paste0(i)) %>%
    relocate(Disease, .after = Exposure)
  
  Index_raw <- full_join(input, next_dis)
}

# Create function to min-max scale within columns
minmax <- function(x, na.rm = TRUE) {
    return((x-min(x))/(max(x)-min(x)))
}

# Pivot input_data wider where Exposure conditions become column names
Index_scale <- pivot_wider(Index_raw, names_from = Exposure, values_from = Index_raw)

# Apply function across Exposure columns, pivot back to long, and join back to input_data
Index_scale_minmax <- Index_scale %>%
  mutate(across("Flaming_eucalyptus":"Smoldering_red_oak", minmax)) %>%
  pivot_longer(!Disease, names_to = "Exposure", values_to = "Index_ExpSc") %>%
  mutate(Coefficient = "Overlap")
Index_input <- full_join(Index_raw, Index_scale_minmax)
```

## Create input dataframe for balloon plot
```{r}
# Assign numeric values to each of your groups to assign the position (x-y coordinates) they will appear on the balloon plot. 
row_df <- data.frame(Disease = c("Asthma","COPD","Hypersensitivity","ILD","Bronchitis","COVID-19","Influenza","Pneumonia"), row = c(8,7,6,5,4,3,2,1))
col_df <- data.frame(Exposure = c("Flaming_eucalyptus","Smoldering_eucalyptus","Flaming_peat","Smoldering_peat","Flaming_pine","Smoldering_pine","Flaming_pine_needles","Smoldering_pine_needles","Flaming_red_oak","Smoldering_red_oak","LPS"), col = c(1,2,3,4,5,6,7,8,9,10,11))

# Join row and col positions to the data frame 
input_data <- Index_input %>%
  left_join(row_df, by = "Disease") %>%
  left_join(col_df, by = "Exposure")

# Manually create axes labels (if you want the labels to be different - I think if you don't do this, it will just keep current names)
# Note that these should be in numeric order for the positions on your axis (row position 1 should have its label in the first slot in the vector)
vars_x_axis <- c("Euc_Fl", "Euc_Sm","Peat_Fl","Peat_Sm","Pine_Fl","Pine_Sm","PineNeedles_Fl","PineNeedles_Sm","RedOak_Fl","RedOak_Sm","LPS")
vars_y_axis <- c("Pneumonia","Influenza","COVID-19","Bronchitis","ILD","Hypersensitivity","COPD","Asthma")
```

## Generate balloon plot
Remember to change ggplot(size=) and geom_text(label=)
```{r}
balloonplot <- ggplot(input_data, aes(x=factor(col), y=factor(row), color=Exposure, size=Index_ExpSc)) +
  scale_fill_brewer(palette="Dark2") +
  geom_point() +    # plot as points
  geom_text(aes(label=formatC(as.numeric(as.character(round(Index_ExpSc, 3))), digits = 3, format = "f"), x=col + 0.25), size=4) +   # display the value next to the "balloons"
  scale_size(range = c(1, 5.5)) + # controls range of dot sizes
  scale_x_discrete(breaks=1:length(vars_x_axis), labels=vars_x_axis, position='top') +   
  scale_y_discrete(breaks=1:length(vars_y_axis), labels=vars_y_axis) +                 
  theme_bw() +
  theme(axis.line = element_blank(),            
        axis.title = element_blank(),          
        panel.border = element_blank(),         
        panel.grid.major.x = element_blank(),   
        panel.grid.minor.x = element_blank(),
        legend.position = "none",
        axis.text = element_text(size = 10, color = "black"),
        axis.ticks = element_blank())
balloonplot
```

## Save balloon plot
Note that the dimensions of the figure you write out will influence how close the rows and columns are to each other. These output dimensions worked well for me for my balloon plot with 4 columns and 5 rows.
NOTE: Export this PDF to TIFF and upload to BioRender for post-processing.
```{r}
pdf(paste0("5_Figures/", date, "_", GeneNum, "_plusLPS_BalloonPlot_Overlap_ExpScaled.pdf"), height = 4, width = 14)
balloonplot
invisible(dev.off())
```

## Second, scale overlap_matrix within DISEASES
```{r}
# Create function to min-max scale within columns
minmax <- function(x, na.rm = TRUE) {
    return((x-min(x))/(max(x)-min(x)))
}

# Pivot input_data wider where Disease conditions become column names
Index_scale <- pivot_wider(Index_raw, names_from = Disease, values_from = Index_raw)

# Apply function across Disease columns, pivot back to long, and join back to input_data
Index_scale_minmax <- Index_scale %>%
  mutate(across("Asthma":"Hypersensitivity", minmax)) %>% #Remember you made this minmax function above
  pivot_longer(!Exposure, names_to = "Disease", values_to = "Index_DisSc") %>%
  mutate(Coefficient = "Overlap")
Index_input <- full_join(Index_input, Index_scale_minmax) 
Index_input <- Index_input %>%
  relocate(Coefficient, .after = Index_DisSc)

# Save as Overlap-specific object
Overlap_input <- Index_input
```

## Create input dataframe for balloon plot
```{r}
# Assign numeric values to each of your groups to assign the position (x-y coordinates) they will appear on the balloon plot. 
row_df <- data.frame(Exposure = c("Flaming_eucalyptus","Smoldering_eucalyptus","Flaming_peat","Smoldering_peat","Flaming_pine","Smoldering_pine","Flaming_pine_needles","Smoldering_pine_needles","Flaming_red_oak","Smoldering_red_oak","LPS"), row = c(11,10,9,8,7,6,5,4,3,2,1))
col_df <- data.frame(Disease = c("Asthma","COPD","Hypersensitivity","ILD","Bronchitis","COVID-19","Influenza","Pneumonia"), col = c(1,2,3,4,5,6,7,8))

# Join row and col positions to the data frame 
input_data <- Index_input %>%
  left_join(row_df, by = "Exposure") %>%
  left_join(col_df, by = "Disease")

# Manually create axes labels (if you want the labels to be different - I think if you don't do this, it will just keep current names)
# Note that these should be in numeric order for the positions on your axis (row position 1 should have its label in the first slot in the vector)
vars_x_axis <- c("Asthma","COPD","Hypersensitivity","ILD","Bronchitis","COVID-19","Influenza","Pneumonia")
vars_y_axis <- c("LPS","RedOak_Sm","RedOak_Fl","PineNeedles_Sm","PineNeedles_Fl","Pine_Sm","Pine_Fl","Peat_Sm","Peat_Fl","LPS","Euc_Sm","Euc_Fl")
```

## Generate balloon plot
Remember to change ggplot(size=) and geom_text(label=)
```{r}
balloonplot <- ggplot(input_data, aes(x=factor(col), y=factor(row), color=Disease, size=Index_DisSc)) +
  scale_fill_brewer(palette="Dark2") +
  geom_point() +    # plot as points
  geom_text(aes(label=formatC(as.numeric(as.character(round(Index_DisSc, 3))), digits = 3, format = "f"), x=col + 0.25), size=4) +   # display the value next to the "balloons"
  scale_size(range = c(1, 5.5)) + # controls range of dot sizes
  scale_x_discrete(breaks=1:length(vars_x_axis), labels=vars_x_axis, position='top') +   
  scale_y_discrete(breaks=1:length(vars_y_axis), labels=vars_y_axis) +                 
  theme_bw() +
  theme(axis.line = element_blank(),            
        axis.title = element_blank(),          
        panel.border = element_blank(),         
        panel.grid.major.x = element_blank(),   
        panel.grid.minor.x = element_blank(),
        legend.position = "none",
        axis.text = element_text(size = 10, color = "black"),
        axis.ticks = element_blank())
balloonplot
```

# Save balloon plot
Note that the dimensions of the figure you write out will influence how close the rows and columns are to each other. These output dimensions worked well for me for my balloon plot with 4 columns and 5 rows.
NOTE: Export this PDF to TIFF and upload to BioRender for post-processing.
```{r}
pdf(paste0("5_Figures/", date, "_", GeneNum, "_plusLPS_BalloonPlot_Overlap_DisScaled.pdf"), height = 6, width = 11)
balloonplot
invisible(dev.off())
```

# DICE-SOERENSEN

## First, scale dice_df within EXPOSURES
```{r}
# Create df to use as basis for bind_rows in for-loop
Index_raw <- dice_df[c(1:11),c(1,13)] #To account for LPS, changed columns to 11 from 10 (Exposures) and rows to 13 from 12 so Asthma is pulled
Index_raw$Disease <- colnames(Index_raw)[2]
Index_raw <- Index_raw %>%
  dplyr::rename("Exposure" = "Condition",
                "Index_raw" = "Asthma") %>%
  relocate(Disease, .after = Exposure)

# Define column names you want looped through (note that Asthma is gone because we defined it in the df above)
dis_cols <- c("Bronchitis","COPD","COVID-19","ILD","Influenza","Pneumonia","Hypersensitivity")
iteration = 13
# Loop through signatures_expo to bind all exposures to main df
for (i in dis_cols) {
  iteration=iteration+1
  input = Index_raw
  
  next_dis <- dice_df[c(1:11),c(1,iteration)] #Changed to 11 from 10 to account for LPS
  next_dis$Disease <- colnames(next_dis)[2]
  next_dis <- next_dis %>%
    dplyr::rename("Exposure" = "Condition",
                  "Index_raw" = paste0(i)) %>%
    relocate(Disease, .after = Exposure)
  
  Index_raw <- full_join(input, next_dis)
}

# Create function to min-max scale within columns
minmax <- function(x, na.rm = TRUE) {
    return((x-min(x))/(max(x)-min(x)))
}

# Pivot input_data wider where Exposure conditions become column names
Index_scale <- pivot_wider(Index_raw, names_from = Exposure, values_from = Index_raw)

# Apply function across Exposure columns, pivot back to long, and join back to input_data
Index_scale_minmax <- Index_scale %>%
  mutate(across("Flaming_eucalyptus":"Smoldering_red_oak", minmax)) %>%
  pivot_longer(!Disease, names_to = "Exposure", values_to = "Index_ExpSc")
Index_input <- full_join(Index_raw, Index_scale_minmax)
```

## Create input dataframe for balloon plot
```{r}
# Assign numeric values to each of your groups to assign the position (x-y coordinates) they will appear on the balloon plot. 
row_df <- data.frame(Disease = c("Asthma","COPD","Hypersensitivity","ILD","Bronchitis","COVID-19","Influenza","Pneumonia"), row = c(8,7,6,5,4,3,2,1))
col_df <- data.frame(Exposure = c("Flaming_eucalyptus","Smoldering_eucalyptus","Flaming_peat","Smoldering_peat","Flaming_pine","Smoldering_pine","Flaming_pine_needles","Smoldering_pine_needles","Flaming_red_oak","Smoldering_red_oak","LPS"), col = c(1,2,3,4,5,6,7,8,9,10,11))

# Join row and col positions to the data frame 
input_data <- Index_input %>%
  left_join(row_df, by = "Disease") %>%
  left_join(col_df, by = "Exposure")

# Manually create axes labels (if you want the labels to be different - I think if you don't do this, it will just keep current names)
# Note that these should be in numeric order for the positions on your axis (row position 1 should have its label in the first slot in the vector)
vars_x_axis <- c("Euc_Fl", "Euc_Sm","Peat_Fl","Peat_Sm","Pine_Fl","Pine_Sm","PineNeedles_Fl","PineNeedles_Sm","RedOak_Fl","RedOak_Sm","LPS")
vars_y_axis <- c("Pneumonia","Influenza","COVID-19","Bronchitis","ILD","Hypersensitivity","COPD","Asthma")
```

## Generate balloon plot
Remember to change ggplot(size=) and geom_text(label=)
```{r}
balloonplot <- ggplot(input_data, aes(x=factor(col), y=factor(row), color=Exposure, size=Index_ExpSc)) +
  scale_fill_brewer(palette="Dark2") +
  geom_point() +    # plot as points
  geom_text(aes(label=formatC(as.numeric(as.character(round(Index_ExpSc, 3))), digits = 3, format = "f"), x=col + 0.25), size=4) +   # display the value next to the "balloons"
  scale_size(range = c(1, 5.5)) + # controls range of dot sizes
  scale_x_discrete(breaks=1:length(vars_x_axis), labels=vars_x_axis, position='top') +   
  scale_y_discrete(breaks=1:length(vars_y_axis), labels=vars_y_axis) +                 
  theme_bw() +
  theme(axis.line = element_blank(),            
        axis.title = element_blank(),          
        panel.border = element_blank(),         
        panel.grid.major.x = element_blank(),   
        panel.grid.minor.x = element_blank(),
        legend.position = "none",
        axis.text = element_text(size = 10, color = "black"),
        axis.ticks = element_blank())
balloonplot
```

## Save balloon plot
Note that the dimensions of the figure you write out will influence how close the rows and columns are to each other. These output dimensions worked well for me for my balloon plot with 4 columns and 5 rows.
NOTE: Export this PDF to TIFF and upload to BioRender for post-processing.
```{r}
pdf(paste0("5_Figures/", date, "_", GeneNum, "_plusLPS_BalloonPlot_Dice_ExpScaled.pdf"), height = 4, width = 14)
balloonplot
invisible(dev.off())
```

## Second, scale overlap_matrix within DISEASES
```{r}
# Create function to min-max scale within columns
minmax <- function(x, na.rm = TRUE) {
    return((x-min(x))/(max(x)-min(x)))
}

# Pivot input_data wider where Disease conditions become column names
Index_scale <- pivot_wider(Index_raw, names_from = Disease, values_from = Index_raw)

# Apply function across Disease columns, pivot back to long, and join back to input_data
Index_scale_minmax <- Index_scale %>%
  mutate(across("Asthma":"Hypersensitivity", minmax)) %>% #Remember you made this minmax function above
  pivot_longer(!Exposure, names_to = "Disease", values_to = "Index_DisSc") %>%
  mutate(Coefficient = "Dice")
Index_input <- full_join(Index_input, Index_scale_minmax) 
Index_input <- Index_input %>%
  relocate(Coefficient, .after = Index_DisSc)

# Save as Dice-specific object
Dice_input <- Index_input
```

## Create input dataframe for balloon plot
```{r}
# Assign numeric values to each of your groups to assign the position (x-y coordinates) they will appear on the balloon plot. 
row_df <- data.frame(Exposure = c("Flaming_eucalyptus","Smoldering_eucalyptus","Flaming_peat","Smoldering_peat","Flaming_pine","Smoldering_pine","Flaming_pine_needles","Smoldering_pine_needles","Flaming_red_oak","Smoldering_red_oak","LPS"), row = c(11,10,9,8,7,6,5,4,3,2,1))
col_df <- data.frame(Disease = c("Asthma","COPD","Hypersensitivity","ILD","Bronchitis","COVID-19","Influenza","Pneumonia"), col = c(1,2,3,4,5,6,7,8))

# Join row and col positions to the data frame 
input_data <- Index_input %>%
  left_join(row_df, by = "Exposure") %>%
  left_join(col_df, by = "Disease")

# Manually create axes labels (if you want the labels to be different - I think if you don't do this, it will just keep current names)
# Note that these should be in numeric order for the positions on your axis (row position 1 should have its label in the first slot in the vector)
vars_x_axis <- c("Asthma","COPD","Hypersensitivity","ILD","Bronchitis","COVID-19","Influenza","Pneumonia")
vars_y_axis <- c("LPS","RedOak_Sm","RedOak_Fl","PineNeedles_Sm","PineNeedles_Fl","Pine_Sm","Pine_Fl","Peat_Sm","Peat_Fl","LPS","Euc_Sm","Euc_Fl")
```

## Generate balloon plot
Remember to change ggplot(size=) and geom_text(label=)
```{r}
balloonplot <- ggplot(input_data, aes(x=factor(col), y=factor(row), color=Disease, size=Index_DisSc)) +
  scale_fill_brewer(palette="Dark2") +
  geom_point() +    # plot as points
  geom_text(aes(label=formatC(as.numeric(as.character(round(Index_DisSc, 3))), digits = 3, format = "f"), x=col + 0.25), size=4) +   # display the value next to the "balloons"
  scale_size(range = c(1, 5.5)) + # controls range of dot sizes
  scale_x_discrete(breaks=1:length(vars_x_axis), labels=vars_x_axis, position='top') +   
  scale_y_discrete(breaks=1:length(vars_y_axis), labels=vars_y_axis) +                 
  theme_bw() +
  theme(axis.line = element_blank(),            
        axis.title = element_blank(),          
        panel.border = element_blank(),         
        panel.grid.major.x = element_blank(),   
        panel.grid.minor.x = element_blank(),
        legend.position = "none",
        axis.text = element_text(size = 10, color = "black"),
        axis.ticks = element_blank())
balloonplot
```

# Save balloon plot
Note that the dimensions of the figure you write out will influence how close the rows and columns are to each other. These output dimensions worked well for me for my balloon plot with 4 columns and 5 rows.
NOTE: Export this PDF to TIFF and upload to BioRender for post-processing.
```{r}
pdf(paste0("5_Figures/", date, "_", GeneNum, "_plusLPS_BalloonPlot_Dice_DisScaled.pdf"), height = 6, width = 11)
balloonplot
invisible(dev.off())
```

# Make master coefficient index df and save
```{r}
# Make master df
Index_input <- Jaccard_input
Index_input <- bind_rows(Index_input, Overlap_input)
Index_input <- bind_rows(Index_input, Dice_input)

# Save onto file generated above
addWorksheet(OUT, "input_values")
writeData(OUT, sheet = "input_values", x = Index_input, rowNames = FALSE)
saveWorkbook(OUT, paste0("4_CountMatrix_Jaccard/", date, "_", GeneNum, "_plusLPS_CountMatrix_MultiMeasures.xlsx"), overwrite = TRUE)
```