---
title: "C_WildfireSim_PostAnalysisNeeds_v2"
author: "Sarah Miller"
date: "2025-05-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(tidyverse)
library(dplyr)
library(janitor)
library(ggplot2)
library(gtools)
library(openxlsx)
library(readr)
#BiocManager::install("clusterProfiler")
library(clusterProfiler)

#date = 250523
#date = 260421
date = 260710
```

# COUNT MATRIX-BASED ANALYSES

## Read-in count matrices
```{r}
# Create list of file names (will use as file path for function below)
input <- list.files(path="4_CountMatrix_Jaccard",
                    pattern="260709_750_plusLPS_CountMatrix_MultiMeasures.xlsx") # Only read-in plusLPS

# Make list of file names and clean so you can use to rename list below
input_names <- gsub("_CountMatrix_MultiMeasures.xlsx","", input)

# Read-in files
setwd("4_CountMatrix_Jaccard")
countmatrices_master <- lapply(input, function(x) {
  read.xlsx(x, sheet = "count_matrix")
})

# Rename list elements with condition
names(countmatrices_master) <- input_names

# Write function to clean column names
clean_CM <- function(data){
  names(data) <- gsub("\\.", "_", names(data)) #Replace periods with underscores in column names
  return(data)
}

# Apply function
countmatrices_master <- lapply(countmatrices_master, clean_CM)
```

# SIMILARITY COEFFICIENT ANALYSES

## Read-in similarity index data
```{r}
# Create list of file names (will use as file path for function below)
input <- list.files(path="4_CountMatrix_Jaccard",
                    pattern="260709_750_plusLPS_CountMatrix_MultiMeasures.xlsx") # Only read-in plusLPS

# Make list of file names and clean so you can use to rename list below
input_names <- gsub("_CountMatrix_MultiMeasures.xlsx","", input)

# Read-in files
setwd("4_CountMatrix_Jaccard")
scores_master <- lapply(input, function(x) {
  read.xlsx(x, sheet = "input_values")
})

# Rename list elements with condition
names(scores_master) <- input_names

# Write function to add GeneNum and Date_filter columns
clean_scores <- function(data_list) {
  result_list <- list() #Make empty list
  for (name in names(data_list)) {
    df <- data_list[[name]] #Extract list element name
    gene_num <- sub(".*?_", "", name) #Extract everything after the first underscore
    df <- df %>% mutate(GeneNum = gene_num) #Add as a column
    date_filter <- sub("_.*", "", name) #Extract everything before the first underscore
    df <- df %>% mutate(Date_filter = date_filter) #Add as a column
    result_list[[name]] <- df #Assign back to list
  }

  return(result_list)
}

# Apply function
scores_master <- clean_scores(scores_master)
```

## Make dataframe of all data (Jaccard, Overlap, Dice; raw, Exp-scaled, Dis-scaled) to make plotting easier

```{r}
# Remove list names so they're not added onto column names
names(scores_master) <- NULL

# Create df to use as basis for bind_rows in for-loop
scores_master_df <- as.data.frame(scores_master[1], row.names = NULL, col.names = names(scores_master[1]))
iteration = 1

# Loop through Disease_master to bind all exposures to main df
for (i in scores_master) {
  iteration=iteration+1
  input = scores_master_df
  
  scores_master_df <- bind_rows(input, as.data.frame(scores_master[iteration], row.names = NULL, col.names = names(scores_master[iteration])))
}
```

## Plot distribution of RAW Jaccard indices

```{r}
# Refactor columns
scores_master_df$GeneNum <- factor(scores_master_df$GeneNum, levels = c("750_plusLPS"))
scores_master_df$Exposure <- factor(scores_master_df$Exposure, levels = c("Flaming_eucalyptus","Smoldering_eucalyptus","Flaming_peat","Smoldering_peat","Flaming_pine","Smoldering_pine","Flaming_pine_needles","Smoldering_pine_needles","Flaming_red_oak","Smoldering_red_oak","LPS"))

# BY EXPOSURE, RAW VALUES
# Jaccard
pJExp <- ggplot(scores_master_df %>% filter(Coefficient == "Jaccard"), aes(x=Exposure, y=Index_raw)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color = Disease), size=2) +
  labs(x = "",
       y = "Jaccard (raw)",
       title = "Distribution of raw Jaccard indices") +
  facet_grid(Date_filter ~ GeneNum) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 70, hjust = 1),
    strip.background = element_rect(fill = "grey25", color = "black", size = 1),  # Change facet label background & add border
    strip.text = element_text(color = "white", face = "bold"),  # Change facet label text color
    panel.border = element_rect(color = "black", fill = NA, size = 1)  # Add black border around facets
  )
pJExp
#ggsave(pJExp, file = paste0("5_Figures/1_ComparisonFigures/", date, "_plusLPS_JaccardRaw.png"), width = 6, height = 5)

# Overlap
pJExp <- ggplot(scores_master_df %>% filter(Coefficient == "Overlap"), aes(x=Exposure, y=Index_raw)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color = Disease), size=2) +
  labs(x = "",
       y = "Overlap (raw)",
       title = "Distribution of raw Overlap indices") +
  facet_grid(Date_filter ~ GeneNum) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 70, hjust = 1),
    strip.background = element_rect(fill = "grey25", color = "black", size = 1),  # Change facet label background & add border
    strip.text = element_text(color = "white", face = "bold"),  # Change facet label text color
    panel.border = element_rect(color = "black", fill = NA, size = 1)  # Add black border around facets
  )
#ggsave(pJExp, file = paste0("5_Figures/1_ComparisonFigures/", date, "_plusLPS_OverlapRaw.png"), width = 6, height = 5)

# Dice
coeff <- "Dice"
pJExp <- ggplot(scores_master_df %>% filter(Coefficient == paste0(coeff)), aes(x=Exposure, y=Index_raw)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color = Disease), size=2) +
  labs(x = "",
       y = paste0(coeff, " (raw)"),
       title = paste0("Distribution of raw ", coeff, " indices")) +
  facet_grid(Date_filter ~ GeneNum) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 70, hjust = 1),
    strip.background = element_rect(fill = "grey25", color = "black", size = 1),  # Change facet label background & add border
    strip.text = element_text(color = "white", face = "bold"),  # Change facet label text color
    panel.border = element_rect(color = "black", fill = NA, size = 1)  # Add black border around facets
  )
#ggsave(pJExp, file = paste0("5_Figures/1_ComparisonFigures/", date, "_plusLPS_", coeff, "Raw.png"), width = 6, height = 5)




# By Disease (Jaccard only)
# RAW
coeff <- "Jaccard"
pDis <- ggplot(scores_master_df %>% filter(Coefficient == paste0(coeff)), aes(x=Disease, y=Index_raw)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color = Exposure), size=2) +
  labs(x = "",
       y = paste0(coeff, " (raw)"),
       title = paste0("Distribution of raw ", coeff, " indices")) +
  facet_grid(Date_filter ~ GeneNum) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.background = element_rect(fill = "grey25", color = "black", size = 1),  # Change facet label background & add border
    strip.text = element_text(color = "white", face = "bold"),  # Change facet label text color
    panel.border = element_rect(color = "black", fill = NA, size = 1)  # Add black border around facets
  )
pDis

# SCALED WITHIN DISEASE
coeff <- "Jaccard"
pDisSc <- ggplot(scores_master_df %>% filter(Coefficient == paste0(coeff)), aes(x=Disease, y=Index_DisSc)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color = Exposure), size=2) +
  labs(x = "",
       y = paste0(coeff, " (scaled with Disease)"),
       title = paste0("Distribution of scaled ", coeff, " indices")) +
  facet_grid(Date_filter ~ GeneNum) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.background = element_rect(fill = "grey25", color = "black", size = 1),  # Change facet label background & add border
    strip.text = element_text(color = "white", face = "bold"),  # Change facet label text color
    panel.border = element_rect(color = "black", fill = NA, size = 1)  # Add black border around facets
  )
pDisSc

# RAW, by GeneNum
coeff <- "Jaccard"
pDis2 <- ggplot(scores_master_df %>% filter(Coefficient == paste0(coeff)), aes(x=Disease, y=Index_raw)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color = GeneNum), size=2) +
  labs(x = "",
       y = "JSI (raw)",
       title = "Distribution of raw Jaccard indices") +
  labs(x = "",
       y = paste0(coeff, " (raw)"),
       title = paste0("Distribution of raw ", coeff, " indices by GeneNum")) +
  facet_wrap(~Exposure, nrow = 2) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.background = element_rect(fill = "grey25", color = "black", size = 1),  # Change facet label background & add border
    strip.text = element_text(color = "white", face = "bold"),  # Change facet label text color
    panel.border = element_rect(color = "black", fill = NA, size = 1)  # Add black border around facets
  )
pDis2
```

## Save plots
```{r}
ggsave(pDis, file = paste0("5_Figures/", date, "_plusLPS_Dis_raw.png"), width = 6, height = 5)
ggsave(pDisSc, file = paste0("5_Figures/", date, "_plusLPS_Dis_scaled.png"), width = 6, height = 5)
ggsave(pDis2, file = paste0("5_Figures/", date, "_plusLPS_Dis_byGeneNum.png"), width = 10, height = 4)
```

## Plot distribution of SCALED Jaccard indices

```{r}
# Create Jaccard-specific dataframe
jaccard_master_df <- scores_master_df %>%
  filter(Coefficient == "Jaccard",
         Date_filter =="260709")

# Refactor columns
jaccard_master_df$GeneNum <- factor(jaccard_master_df$GeneNum, levels = c("750_plusLPS"))
jaccard_master_df$Exposure <- factor(jaccard_master_df$Exposure, levels = c("Flaming_eucalyptus","Smoldering_eucalyptus","Flaming_peat","Smoldering_peat","Flaming_pine","Smoldering_pine","Flaming_pine_needles","Smoldering_pine_needles","Flaming_red_oak","Smoldering_red_oak","LPS"))

# By Exposure
pExp <- ggplot(jaccard_master_df, aes(x=Exposure, y=Index_raw)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color = Disease), size=2) +
  labs(x = "",
       y = "JSI (raw)",
       title = "Distribution of raw Jaccard indices") +
  facet_wrap(~GeneNum, nrow = 1) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.background = element_rect(fill = "grey25", color = "black", size = 1),  # Change facet label background & add border
    strip.text = element_text(color = "white", face = "bold"),  # Change facet label text color
    panel.border = element_rect(color = "black", fill = NA, size = 1)  # Add black border around facets
  )
pExp

pExp2 <- ggplot(jaccard_master_df, aes(x=Exposure, y=Index_raw)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color = GeneNum), size=2) +
  labs(x = "",
       y = "JSI (raw)",
       title = "Distribution of raw Jaccard indices") +
  facet_wrap(~Disease, nrow = 2) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.background = element_rect(fill = "grey25", color = "black", size = 1),  # Change facet label background & add border
    strip.text = element_text(color = "white", face = "bold"),  # Change facet label text color
    panel.border = element_rect(color = "black", fill = NA, size = 1)  # Add black border around facets
  )
pExp2

# By Disease
pDis <- ggplot(jaccard_master_df, aes(x=Disease, y=Index_raw)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color = Exposure), size=2) +
  labs(x = "",
       y = "JSI (raw)",
       title = "Distribution of raw Jaccard indices") +
  facet_wrap(~GeneNum, nrow = 1) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.background = element_rect(fill = "grey25", color = "black", size = 1),  # Change facet label background & add border
    strip.text = element_text(color = "white", face = "bold"),  # Change facet label text color
    panel.border = element_rect(color = "black", fill = NA, size = 1)  # Add black border around facets
  )
pDis

pDis2 <- ggplot(jaccard_master_df, aes(x=Disease, y=Index_raw)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color = GeneNum), size=2) +
  labs(x = "",
       y = "JSI (raw)",
       title = "Distribution of raw Jaccard indices") +
  facet_wrap(~Exposure, nrow = 2) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.background = element_rect(fill = "grey25", color = "black", size = 1),  # Change facet label background & add border
    strip.text = element_text(color = "white", face = "bold"),  # Change facet label text color
    panel.border = element_rect(color = "black", fill = NA, size = 1)  # Add black border around facets
  )
pDis2
```

## Save plots
```{r}
ggsave(pExp, file = paste0("5_Figures/", date, "_plusLPS_Exp_byGeneNum.png"), width = 10, height = 4)
ggsave(pExp2, file = paste0("5_Figures/", date, "_plusLPS_Exp_byDis.png"), width = 8, height = 6)
ggsave(pDis, file = paste0("5_Figures/", date, "_plusLPS_Dis_byGeneNum.png"), width = 10, height = 4)
ggsave(pDis2, file = paste0("5_Figures/", date, "_plusLPS_Dis_byExp.png"), width = 8, height = 6)
```


# Generate histograms of highly shared genes
```{r}
# Write function to generate sum column
clean_CM <- function(data){
  data$sum <- rowSums(data[,2:ncol(data)]) #Sum across rows
  data$Exposure_sum <- rowSums(data[,2:12]) #Sum across Exposure rows
  data$Disease_sum <- rowSums(data[,13:20]) #Sum across Disease rows
  data <- arrange(data, desc(sum)) #Order by descending sum so highly shared genes are at top
  return(data)
}

# Apply function
countmatrices_master <- lapply(countmatrices_master, clean_CM)
#countmatrices_master <- lapply(countmatrices_master, clean_scores)

# Assign INPUT to be used in histogram
INPUT <- countmatrices_master[[1]]
GeneNum = "750_plusLPS"
#write_csv(INPUT, file = "4_CountMatrix_Jaccard/test.csv")

# Create table to use as input for histogram
# Define the range of numbers to count (1 through whatever is the max "sum" column value, e.g. highest number of exposures over which a given gene is shared)
numbers_to_count <- 1:max(INPUT$sum, na.rm = TRUE)
# Count occurrences of each number in the column
count_table <- table(factor(INPUT$sum, levels = numbers_to_count))
# Convert the count table to a data frame
result_df <- data.frame(Number = as.integer(names(count_table)), Count = as.integer(count_table))

# Define color sequence for histogram and define Number as characters
#result_df$Color <- scales::col_numeric(palette = "Greens", domain = range(result_df$Count))(result_df$Count)
result_df$Color <- scales::col_numeric(palette = "Purples", domain = range(result_df$Count))(result_df$Count)

# Plot histogram
pHist <-ggplot(result_df, aes(x=factor(Number), y=Count, fill=Color)) +
  geom_col(color="black") +
  scale_fill_identity() +
  geom_text(aes(label = Count), vjust = -1, colour = "black") +
  xlab("Number") +
  labs(x = "Number of conditions where genes are shared",
       y = "Number of shared genes") +
  ylim(c(0, 2600)) +
  theme_minimal() +
  theme(
    strip.background = element_rect(fill = "grey25", color = "black", size = 1),  # Change facet label background & add border
    strip.text = element_text(color = "white", face = "bold"),  # Change facet label text color
    panel.border = element_rect(color = "black", fill = NA, size = 1)  # Add black border around facets
  )
pHist

# Save plot
#ggsave(pHist, file = "5_Figures/250523_750_Histogram.png", width = 6, height = 4)
ggsave(pHist, file = paste0("5_Figures/", date, "_750_plusLPS_Histogram_purple.png"), width = 6, height = 4)
```

# Create exposure vs disease distribution tables and plot
```{r}
# Write function
make_distrib_table <- function(data){
  # Create Exposure sum table
  exp <- data[,c(1,21,22)]
  exp <- exp %>%
    mutate(Condition = "Exposure") %>%
    dplyr::rename(Condition_counts = Exposure_sum)
  # Create Disease sum table
  dis <- data[,c(1,21,23)]
  dis <- dis %>%
    mutate(Condition = "Disease") %>%
    dplyr::rename(Condition_counts = Disease_sum)
  # Join together and arrange by Human_gene and descending sum column
  data_newtable <- full_join(exp, dis)
  data_newtable <- data_newtable %>%
    group_by(Human_gene) %>%
    arrange(desc(sum))
  
  return(data_newtable)
}

# Apply function
condition_distrib <- lapply(countmatrices_master, make_distrib_table)


# Assign INPUT
INPUT <- condition_distrib[[1]]
GeneNum = "750_plusLPS"
# Make Human_gene a factor with levels in the order of appearance
INPUT$Human_gene <- factor(INPUT$Human_gene, levels = unique(INPUT$Human_gene))

# Plot genes of interest distribution by condition
p1 <- ggplot(INPUT %>% filter(sum > 9), aes(x = Human_gene, y = Condition_counts, fill = Condition)) +
  geom_bar(color = "black", stat = "identity") +
  #scale_fill_manual(values = c("Exposure" = "orange2", "Disease" = "lightblue")) +
  scale_fill_manual(values = c("Exposure" = "#94221F", "Disease" = "#3E7274")) +
  labs(x = "", y = "Number of conditions", fill = "Condition") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        axis.text.y = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10))
p1

# Save plot
# width = 14 for 750_plusLPS
ggsave(p1, file = paste0("5_Figures/", date, "_750_plusLPS_ConditionDistributions_redblue.png"), width = 14, height = 4)
```

# Plot genes of interest where at least 1 disease is associated
```{r}
# Assign INPUT
INPUT <- condition_distrib[[1]]
GeneNum = "750_plusLPS"
# Make Human_gene a factor with levels in the order of appearance
INPUT$Human_gene <- factor(INPUT$Human_gene, levels = unique(INPUT$Human_gene))
INPUT <- INPUT %>%
  group_by(Human_gene) %>%
  filter(sum > 9,
         (any(Condition == "Disease" & Condition_counts > 0)))

TGCs_10cond1dis <- INPUT %>%
  filter(duplicated(Human_gene))

# Plot genes of interest distribution by condition
p1 <- ggplot(INPUT, aes(x = Human_gene, y = Condition_counts, fill = Condition)) +
  geom_bar(color = "black", stat = "identity") +
  #scale_fill_manual(values = c("Exposure" = "orange2", "Disease" = "lightblue")) +
  scale_fill_manual(values = c("Exposure" = "#94221F", "Disease" = "#3E7274")) +
  labs(x = "", y = "Number of conditions", fill = "Condition") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        axis.text.y = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10))
p1

# Save plot
# width = 14 for 750_plusLPS
ggsave(p1, file = paste0("5_Figures/", date, "_750_plusLPS_ConditionDistributions_AtLeast1Dis_redblue.png"), width = 14, height = 4)
```

## Filter lists for genes of interest where at least 1 disease contributes
```{r}
# Write function
filter_CM <- function(data){
  data <- data %>%
    group_by(Human_gene) %>%
    filter(sum > 9,
           Disease_sum > 0)
  
  return(data)
}
# Apply function
countmatrices_10cond1dis <- lapply(countmatrices_master, filter_CM)

# Remove non-250523 matrices
countmatrices_10cond1dis <- countmatrices_10cond1dis[-c(1,2)]
```

## Identify which disease contributes the most to genes of interest
```{r}
# Write function
disease_contributor <- function(data){
  col_sums <- colSums(data[ , 2:19], na.rm = TRUE)
  
  return(col_sums)
}
# Apply function
disease_contributions <- lapply(countmatrices_10cond1dis, disease_contributor)
```

# Identify which conditions (particularly diseases) contribute the most toward each signature
```{r}
# Assign INPUT. [1] = 1000, [2] = 500, [3] = 750
INPUT <- data.frame(disease_contributions[3])
colnames(INPUT) <- c("Gene_counts")
INPUT <- INPUT %>%
  rownames_to_column("Condition")
INPUT$Condition_type <- NA
INPUT$Condition_type[1:10] <- "Exposure"
INPUT$Condition_type[11:18] <- "Disease"

# Make Condition a factor with levels in the order of descending Gene_counts
INPUT <- INPUT %>%
  mutate(Condition = factor(Condition, levels = unique(Condition[order(-Gene_counts)])))

# Plot genes of interest distribution by condition
p1 <- ggplot(INPUT, aes(x = Condition, y = Gene_counts, fill = Condition_type)) +
  geom_bar(color = "black", stat = "identity") +
  #scale_fill_manual(values = c("Exposure" = "orange2", "Disease" = "lightblue")) +
  scale_fill_manual(values = c("Exposure" = "#94221F", "Disease" = "#3E7274")) +
  labs(x = "", y = "Genes of interest", fill = "Condition_type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        axis.text.y = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10))
p1

# Save plot
ggsave(p1, file = "5_Figures/250523_750_ConditionDistributions_AtLeast1Dis_TopContrib_redblue.png", width = 6, height = 4)
```

## Convert countmatrices_10cond1dis to condition-specific lists so they can be read into similarity coefficient functions below.
```{r}
# countmatrices_10cond1dis: [1] = 1000, [2] = 500, [3] = 750
INPUT <- countmatrices_10cond1dis[2]
filters <- "10Cond1Dis"

# Clean
list_name <- names(INPUT) #Save list name so you know what you're working with and can paste below
names(INPUT) <- NULL #Remove list name so it's not appended to every column name when turned into df
INPUT <- data.frame(INPUT) %>%
  select(-c(sum, Exposure_sum, Disease_sum)) #Remove columns we don't need
names(INPUT)[15] <- "COVID-19" #Rename COVID-19 column, it replaced dash with period

# Get condition column names (excluding the Human_gene column)
condition_cols <- setdiff(names(INPUT), "Human_gene")
# Create list of dataframes, one per condition
INPUT_lists <- lapply(condition_cols, function(cond) {
  INPUT[INPUT[[cond]] == 1, "Human_gene", drop = FALSE]
})
# Name each list element with the condition name
names(INPUT_lists) <- condition_cols
```

## Calculate similarity coefficients using our filtered genes of interest
```{r}
# Define the three similarity functions
jaccard_similarity <- function(set1, set2) {
  intersect_len <- length(intersect(set1, set2))
  union_len <- length(union(set1, set2))
  return(intersect_len / union_len)
}

dice_coefficient <- function(set1, set2) {
  intersect_len <- length(intersect(set1, set2))
  return(2 * intersect_len / (length(set1) + length(set2)))
}

overlap_coefficient <- function(set1, set2) {
  intersect_len <- length(intersect(set1, set2))
  return(intersect_len / min(length(set1), length(set2)))
}

# Extract sets of genes
gene_sets <- lapply(INPUT_lists, function(df) unique(df$Human_gene))

# Number of signatures
n <- length(gene_sets)
labels <- names(INPUT_lists)

# Initialize matrices
jaccard_matrix <- matrix(0, n, n, dimnames = list(labels, labels))
dice_matrix <- matrix(0, n, n, dimnames = list(labels, labels))
overlap_matrix <- matrix(0, n, n, dimnames = list(labels, labels))

# Compute pairwise similarities
for (i in seq_len(n)) {
  for (j in seq_len(n)) {
    set1 <- gene_sets[[i]]
    set2 <- gene_sets[[j]]
    
    jaccard_matrix[i, j] <- jaccard_similarity(set1, set2)
    dice_matrix[i, j] <- dice_coefficient(set1, set2)
    overlap_matrix[i, j] <- overlap_coefficient(set1, set2)
  }
}

# Convert to data frames 
jaccard_df <- as.data.frame(jaccard_matrix) %>%
  rownames_to_column("Condition")
dice_df <- as.data.frame(dice_matrix) %>%
  rownames_to_column("Condition")
overlap_df <- as.data.frame(overlap_matrix) %>%
  rownames_to_column("Condition")
```

## Create and save Excel of all data
```{r}
# Create Excel workbook
OUT <- createWorkbook()
# Add Excel sheets and write-out both filtered count matrix and similarity matrices
addWorksheet(OUT, "count_matrix_filtered")
writeData(OUT, sheet = "count_matrix_filtered", x = INPUT)
addWorksheet(OUT, "jaccard_matrix")
writeData(OUT, sheet = "jaccard_matrix", x = jaccard_df)
addWorksheet(OUT, "overlap_matrix")
writeData(OUT, sheet = "overlap_matrix", x = overlap_df)
addWorksheet(OUT, "dice_matrix")
writeData(OUT, sheet = "dice_matrix", x = dice_df)
# Export file
saveWorkbook(OUT, paste0("4_CountMatrix_Jaccard/", list_name, "_CountMatrix_MultiMeasures_", filters, ".xlsx"), overwrite = TRUE)
```

# GENERATE BALLOON PLOTS

# JACCARD

## First, scale jaccard_df within EXPOSURES
```{r}
# Create df to use as basis for bind_rows in for-loop
Index_raw <- jaccard_df[c(1:10),c(1,12)]
Index_raw$Disease <- colnames(Index_raw)[2]
Index_raw <- Index_raw %>%
  rename("Exposure" = "Condition",
         "Index_raw" = "Asthma") %>%
  relocate(Disease, .after = Exposure)

# Define column names you want looped through (note that Asthma is gone because we defined it in the df above)
dis_cols <- c("Bronchitis","COPD","COVID-19","ILD","Influenza","Pneumonia","Hypersensitivity")
iteration = 12
# Loop through signatures_expo to bind all exposures to main df
for (i in dis_cols) {
  iteration=iteration+1
  input = Index_raw
  
  next_dis <- jaccard_df[c(1:10),c(1,iteration)]
  next_dis$Disease <- colnames(next_dis)[2]
  next_dis <- next_dis %>%
    rename("Exposure" = "Condition",
           "Index_raw" = paste0(i)) %>%
    relocate(Disease, .after = Exposure)
  
  Index_raw <- full_join(input, next_dis)
}

# Create function to min-max scale within columns
minmax <- function(x, na.rm = TRUE) {
    return((x-min(x))/(max(x)-min(x)))
}

# Pivot input_data wider where Exposure conditions become column names
Index_scale <- pivot_wider(Index_raw, names_from = Exposure, values_from = Index_raw)

# Apply function across Exposure columns, pivot back to long, and join back to input_data
Index_scale_minmax <- Index_scale %>%
  mutate(across("Flaming_eucalyptus":"Smoldering_red_oak", minmax)) %>%
  pivot_longer(!Disease, names_to = "Exposure", values_to = "Index_ExpSc")
Index_input <- full_join(Index_raw, Index_scale_minmax)
```

## Create input dataframe for balloon plot
```{r}
# Assign numeric values to each of your groups to assign the position (x-y coordinates) they will appear on the balloon plot. 
row_df <- data.frame(Disease = c("Asthma","COPD","Hypersensitivity","ILD","Bronchitis","COVID-19","Influenza","Pneumonia"), row = c(8,7,6,5,4,3,2,1))
col_df <- data.frame(Exposure = c("Flaming_eucalyptus", "Smoldering_eucalyptus","Flaming_peat","Smoldering_peat","Flaming_pine","Smoldering_pine","Flaming_pine_needles","Smoldering_pine_needles","Flaming_red_oak","Smoldering_red_oak"), col = c(1,2,3,4,5,6,7,8,9,10))

# Join row and col positions to the data frame 
input_data <- Index_input %>%
  left_join(row_df, by = "Disease") %>%
  left_join(col_df, by = "Exposure")

# Manually create axes labels (if you want the labels to be different - I think if you don't do this, it will just keep current names)
# Note that these should be in numeric order for the positions on your axis (row position 1 should have its label in the first slot in the vector)
vars_x_axis <- c("Euc_Fl", "Euc_Sm","Peat_Fl","Peat_Sm","Pine_Fl","Pine_Sm","PineNeedles_Fl","PineNeedles_Sm","RedOak_Fl","RedOak_Sm")
vars_y_axis <- c("Pneumonia","Influenza","COVID-19","Bronchitis","ILD","Hypersensitivity","COPD","Asthma")
```

## Generate balloon plot
Remember to change ggplot(size=) and geom_text(label=)
```{r}
balloonplot <- ggplot(input_data, aes(x=factor(col), y=factor(row), color=Exposure, size=Index_ExpSc)) +
  scale_fill_brewer(palette="Dark2") +
  geom_point() +    # plot as points
  geom_text(aes(label=formatC(as.numeric(as.character(round(Index_ExpSc, 3))), digits = 3, format = "f"), x=col + 0.25), size=4) +   # display the value next to the "balloons"
  scale_size(range = c(1, 5.5)) + # controls range of dot sizes
  scale_x_discrete(breaks=1:length(vars_x_axis), labels=vars_x_axis, position='top') +   
  scale_y_discrete(breaks=1:length(vars_y_axis), labels=vars_y_axis) +                 
  theme_bw() +
  theme(axis.line = element_blank(),            
        axis.title = element_blank(),          
        panel.border = element_blank(),         
        panel.grid.major.x = element_blank(),   
        panel.grid.minor.x = element_blank(),
        legend.position = "none",
        axis.text = element_text(size = 10, color = "black"),
        axis.ticks = element_blank())
balloonplot
```

## Save balloon plot
Note that the dimensions of the figure you write out will influence how close the rows and columns are to each other. These output dimensions worked well for me for my balloon plot with 4 columns and 5 rows.
NOTE: Export this PDF to TIFF and upload to BioRender for post-processing.
```{r}
pdf(paste0("5_Figures/", list_name, "_BalloonPlot_Jaccard_ExpScaled_", filters, ".pdf"), height = 4, width = 13)
balloonplot
invisible(dev.off())
```

## Second, scale JSIs within DISEASES
```{r}
# Create function to min-max scale within columns
minmax <- function(x, na.rm = TRUE) {
    return((x-min(x))/(max(x)-min(x)))
}

# Pivot input_data wider where Disease conditions become column names
Index_scale <- pivot_wider(Index_raw, names_from = Disease, values_from = Index_raw)

# Apply function across Disease columns, pivot back to long, and join back to input_data
Index_scale_minmax <- Index_scale %>%
  mutate(across("Asthma":"Hypersensitivity", minmax)) %>% #Remember you made this minmax function above
  pivot_longer(!Exposure, names_to = "Disease", values_to = "Index_DisSc")
Index_input <- full_join(Index_input, Index_scale_minmax) %>%
  mutate(Coefficient = "Jaccard")

# Save as Jaccard-specific object
Jaccard_input <- Index_input
```

## Create input dataframe for balloon plot
```{r}
# Assign numeric values to each of your groups to assign the position (x-y coordinates) they will appear on the balloon plot. 
row_df <- data.frame(Exposure = c("Flaming_eucalyptus", "Smoldering_eucalyptus","Flaming_peat","Smoldering_peat","Flaming_pine","Smoldering_pine","Flaming_pine_needles","Smoldering_pine_needles","Flaming_red_oak","Smoldering_red_oak"), row = c(10,9,8,7,6,5,4,3,2,1))
col_df <- data.frame(Disease = c("Asthma","COPD","Hypersensitivity","ILD","Bronchitis","COVID-19","Influenza","Pneumonia"), col = c(1,2,3,4,5,6,7,8))

# Join row and col positions to the data frame 
input_data <- Index_input %>%
  left_join(row_df, by = "Exposure") %>%
  left_join(col_df, by = "Disease")

# Manually create axes labels (if you want the labels to be different - I think if you don't do this, it will just keep current names)
# Note that these should be in numeric order for the positions on your axis (row position 1 should have its label in the first slot in the vector)
vars_x_axis <- c("Asthma","COPD","Hypersensitivity","ILD","Bronchitis","COVID-19","Influenza","Pneumonia")
vars_y_axis <- c("RedOak_Sm","RedOak_Fl","PineNeedles_Sm","PineNeedles_Fl","Pine_Sm","Pine_Fl","Peat_Sm","Peat_Fl","Euc_Sm","Euc_Fl")
```

## Generate balloon plot
Remember to change ggplot(size=) and geom_text(label=)
```{r}
balloonplot <- ggplot(input_data, aes(x=factor(col), y=factor(row), color=Disease, size=Index_DisSc)) +
  scale_fill_brewer(palette="Dark2") +
  geom_point() +    # plot as points
  geom_text(aes(label=formatC(as.numeric(as.character(round(Index_DisSc, 3))), digits = 3, format = "f"), x=col + 0.25), size=4) +   # display the value next to the "balloons"
  scale_size(range = c(1, 5.5)) + # controls range of dot sizes
  scale_x_discrete(breaks=1:length(vars_x_axis), labels=vars_x_axis, position='top') +   
  scale_y_discrete(breaks=1:length(vars_y_axis), labels=vars_y_axis) +                 
  theme_bw() +
  theme(axis.line = element_blank(),            
        axis.title = element_blank(),          
        panel.border = element_blank(),         
        panel.grid.major.x = element_blank(),   
        panel.grid.minor.x = element_blank(),
        legend.position = "none",
        axis.text = element_text(size = 10, color = "black"),
        axis.ticks = element_blank())
balloonplot
```

# Save balloon plot
Note that the dimensions of the figure you write out will influence how close the rows and columns are to each other. These output dimensions worked well for me for my balloon plot with 4 columns and 5 rows.
NOTE: Export this PDF to TIFF and upload to BioRender for post-processing.
```{r}
pdf(paste0("5_Figures/", list_name, "_BalloonPlot_Jaccard_DisScaled_", filters, ".pdf"), height = 5, width = 11)
balloonplot
invisible(dev.off())
```

# OVERLAP

## First, scale overlap_df within EXPOSURES
```{r}
# Create df to use as basis for bind_rows in for-loop
Index_raw <- overlap_df[c(1:10),c(1,12)]
Index_raw$Disease <- colnames(Index_raw)[2]
Index_raw <- Index_raw %>%
  rename("Exposure" = "Condition",
         "Index_raw" = "Asthma") %>%
  relocate(Disease, .after = Exposure)

# Define column names you want looped through (note that Asthma is gone because we defined it in the df above)
dis_cols <- c("Bronchitis","COPD","COVID-19","ILD","Influenza","Pneumonia","Hypersensitivity")
iteration = 12
# Loop through signatures_expo to bind all exposures to main df
for (i in dis_cols) {
  iteration=iteration+1
  input = Index_raw
  
  next_dis <- overlap_df[c(1:10),c(1,iteration)]
  next_dis$Disease <- colnames(next_dis)[2]
  next_dis <- next_dis %>%
    rename("Exposure" = "Condition",
           "Index_raw" = paste0(i)) %>%
    relocate(Disease, .after = Exposure)
  
  Index_raw <- full_join(input, next_dis)
}

# Create function to min-max scale within columns
minmax <- function(x, na.rm = TRUE) {
    return((x-min(x))/(max(x)-min(x)))
}

# Pivot input_data wider where Exposure conditions become column names
Index_scale <- pivot_wider(Index_raw, names_from = Exposure, values_from = Index_raw)

# Apply function across Exposure columns, pivot back to long, and join back to input_data
Index_scale_minmax <- Index_scale %>%
  mutate(across("Flaming_eucalyptus":"Smoldering_red_oak", minmax)) %>%
  pivot_longer(!Disease, names_to = "Exposure", values_to = "Index_ExpSc") %>%
  mutate(Coefficient = "Overlap")
Index_input <- full_join(Index_raw, Index_scale_minmax)
```

## Create input dataframe for balloon plot
```{r}
# Assign numeric values to each of your groups to assign the position (x-y coordinates) they will appear on the balloon plot. 
row_df <- data.frame(Disease = c("Asthma","COPD","Hypersensitivity","ILD","Bronchitis","COVID-19","Influenza","Pneumonia"), row = c(8,7,6,5,4,3,2,1))
col_df <- data.frame(Exposure = c("Flaming_eucalyptus", "Smoldering_eucalyptus","Flaming_peat","Smoldering_peat","Flaming_pine","Smoldering_pine","Flaming_pine_needles","Smoldering_pine_needles","Flaming_red_oak","Smoldering_red_oak"), col = c(1,2,3,4,5,6,7,8,9,10))

# Join row and col positions to the data frame 
input_data <- Index_input %>%
  left_join(row_df, by = "Disease") %>%
  left_join(col_df, by = "Exposure")

# Manually create axes labels (if you want the labels to be different - I think if you don't do this, it will just keep current names)
# Note that these should be in numeric order for the positions on your axis (row position 1 should have its label in the first slot in the vector)
vars_x_axis <- c("Euc_Fl", "Euc_Sm","Peat_Fl","Peat_Sm","Pine_Fl","Pine_Sm","PineNeedles_Fl","PineNeedles_Sm","RedOak_Fl","RedOak_Sm")
vars_y_axis <- c("Pneumonia","Influenza","COVID-19","Bronchitis","ILD","Hypersensitivity","COPD","Asthma")
```

## Generate balloon plot
Remember to change ggplot(size=) and geom_text(label=)
```{r}
balloonplot <- ggplot(input_data, aes(x=factor(col), y=factor(row), color=Exposure, size=Index_ExpSc)) +
  scale_fill_brewer(palette="Dark2") +
  geom_point() +    # plot as points
  geom_text(aes(label=formatC(as.numeric(as.character(round(Index_ExpSc, 3))), digits = 3, format = "f"), x=col + 0.25), size=4) +   # display the value next to the "balloons"
  scale_size(range = c(1, 5.5)) + # controls range of dot sizes
  scale_x_discrete(breaks=1:length(vars_x_axis), labels=vars_x_axis, position='top') +   
  scale_y_discrete(breaks=1:length(vars_y_axis), labels=vars_y_axis) +                 
  theme_bw() +
  theme(axis.line = element_blank(),            
        axis.title = element_blank(),          
        panel.border = element_blank(),         
        panel.grid.major.x = element_blank(),   
        panel.grid.minor.x = element_blank(),
        legend.position = "none",
        axis.text = element_text(size = 10, color = "black"),
        axis.ticks = element_blank())
balloonplot
```

## Save balloon plot
Note that the dimensions of the figure you write out will influence how close the rows and columns are to each other. These output dimensions worked well for me for my balloon plot with 4 columns and 5 rows.
NOTE: Export this PDF to TIFF and upload to BioRender for post-processing.
```{r}
pdf(paste0("5_Figures/", list_name, "_BalloonPlot_Overlap_ExpScaled_", filters, ".pdf"), height = 4, width = 13)
balloonplot
invisible(dev.off())
```

## Second, scale overlap_matrix within DISEASES
```{r}
# Create function to min-max scale within columns
minmax <- function(x, na.rm = TRUE) {
    return((x-min(x))/(max(x)-min(x)))
}

# Pivot input_data wider where Disease conditions become column names
Index_scale <- pivot_wider(Index_raw, names_from = Disease, values_from = Index_raw)

# Apply function across Disease columns, pivot back to long, and join back to input_data
Index_scale_minmax <- Index_scale %>%
  mutate(across("Asthma":"Hypersensitivity", minmax)) %>% #Remember you made this minmax function above
  pivot_longer(!Exposure, names_to = "Disease", values_to = "Index_DisSc") %>%
  mutate(Coefficient = "Overlap")
Index_input <- full_join(Index_input, Index_scale_minmax) 
Index_input <- Index_input %>%
  relocate(Coefficient, .after = Index_DisSc)

# Save as Overlap-specific object
Overlap_input <- Index_input
```

## Create input dataframe for balloon plot
```{r}
# Assign numeric values to each of your groups to assign the position (x-y coordinates) they will appear on the balloon plot. 
row_df <- data.frame(Exposure = c("Flaming_eucalyptus", "Smoldering_eucalyptus","Flaming_peat","Smoldering_peat","Flaming_pine","Smoldering_pine","Flaming_pine_needles","Smoldering_pine_needles","Flaming_red_oak","Smoldering_red_oak"), row = c(10,9,8,7,6,5,4,3,2,1))
col_df <- data.frame(Disease = c("Asthma","COPD","Hypersensitivity","ILD","Bronchitis","COVID-19","Influenza","Pneumonia"), col = c(1,2,3,4,5,6,7,8))

# Join row and col positions to the data frame 
input_data <- Index_input %>%
  left_join(row_df, by = "Exposure") %>%
  left_join(col_df, by = "Disease")

# Manually create axes labels (if you want the labels to be different - I think if you don't do this, it will just keep current names)
# Note that these should be in numeric order for the positions on your axis (row position 1 should have its label in the first slot in the vector)
vars_x_axis <- c("Asthma","COPD","Hypersensitivity","ILD","Bronchitis","COVID-19","Influenza","Pneumonia")
vars_y_axis <- c("RedOak_Sm","RedOak_Fl","PineNeedles_Sm","PineNeedles_Fl","Pine_Sm","Pine_Fl","Peat_Sm","Peat_Fl","Euc_Sm","Euc_Fl")
```

## Generate balloon plot
Remember to change ggplot(size=) and geom_text(label=)
```{r}
balloonplot <- ggplot(input_data, aes(x=factor(col), y=factor(row), color=Disease, size=Index_DisSc)) +
  scale_fill_brewer(palette="Dark2") +
  geom_point() +    # plot as points
  geom_text(aes(label=formatC(as.numeric(as.character(round(Index_DisSc, 3))), digits = 3, format = "f"), x=col + 0.25), size=4) +   # display the value next to the "balloons"
  scale_size(range = c(1, 5.5)) + # controls range of dot sizes
  scale_x_discrete(breaks=1:length(vars_x_axis), labels=vars_x_axis, position='top') +   
  scale_y_discrete(breaks=1:length(vars_y_axis), labels=vars_y_axis) +                 
  theme_bw() +
  theme(axis.line = element_blank(),            
        axis.title = element_blank(),          
        panel.border = element_blank(),         
        panel.grid.major.x = element_blank(),   
        panel.grid.minor.x = element_blank(),
        legend.position = "none",
        axis.text = element_text(size = 10, color = "black"),
        axis.ticks = element_blank())
balloonplot
```

# Save balloon plot
Note that the dimensions of the figure you write out will influence how close the rows and columns are to each other. These output dimensions worked well for me for my balloon plot with 4 columns and 5 rows.
NOTE: Export this PDF to TIFF and upload to BioRender for post-processing.
```{r}
pdf(paste0("5_Figures/", list_name, "_BalloonPlot_Overlap_DisScaled_", filters, ".pdf"), height = 5, width = 11)
balloonplot
invisible(dev.off())
```

# DICE-SOERENSEN

## First, scale dice_df within EXPOSURES
```{r}
# Create df to use as basis for bind_rows in for-loop
Index_raw <- dice_df[c(1:10),c(1,12)]
Index_raw$Disease <- colnames(Index_raw)[2]
Index_raw <- Index_raw %>%
  rename("Exposure" = "Condition",
         "Index_raw" = "Asthma") %>%
  relocate(Disease, .after = Exposure)

# Define column names you want looped through (note that Asthma is gone because we defined it in the df above)
dis_cols <- c("Bronchitis","COPD","COVID-19","ILD","Influenza","Pneumonia","Hypersensitivity")
iteration = 12
# Loop through signatures_expo to bind all exposures to main df
for (i in dis_cols) {
  iteration=iteration+1
  input = Index_raw
  
  next_dis <- dice_df[c(1:10),c(1,iteration)]
  next_dis$Disease <- colnames(next_dis)[2]
  next_dis <- next_dis %>%
    rename("Exposure" = "Condition",
           "Index_raw" = paste0(i)) %>%
    relocate(Disease, .after = Exposure)
  
  Index_raw <- full_join(input, next_dis)
}

# Create function to min-max scale within columns
minmax <- function(x, na.rm = TRUE) {
    return((x-min(x))/(max(x)-min(x)))
}

# Pivot input_data wider where Exposure conditions become column names
Index_scale <- pivot_wider(Index_raw, names_from = Exposure, values_from = Index_raw)

# Apply function across Exposure columns, pivot back to long, and join back to input_data
Index_scale_minmax <- Index_scale %>%
  mutate(across("Flaming_eucalyptus":"Smoldering_red_oak", minmax)) %>%
  pivot_longer(!Disease, names_to = "Exposure", values_to = "Index_ExpSc")
Index_input <- full_join(Index_raw, Index_scale_minmax)
```

## Create input dataframe for balloon plot
```{r}
# Assign numeric values to each of your groups to assign the position (x-y coordinates) they will appear on the balloon plot. 
row_df <- data.frame(Disease = c("Asthma","COPD","Hypersensitivity","ILD","Bronchitis","COVID-19","Influenza","Pneumonia"), row = c(8,7,6,5,4,3,2,1))
col_df <- data.frame(Exposure = c("Flaming_eucalyptus", "Smoldering_eucalyptus","Flaming_peat","Smoldering_peat","Flaming_pine","Smoldering_pine","Flaming_pine_needles","Smoldering_pine_needles","Flaming_red_oak","Smoldering_red_oak"), col = c(1,2,3,4,5,6,7,8,9,10))

# Join row and col positions to the data frame 
input_data <- Index_input %>%
  left_join(row_df, by = "Disease") %>%
  left_join(col_df, by = "Exposure")

# Manually create axes labels (if you want the labels to be different - I think if you don't do this, it will just keep current names)
# Note that these should be in numeric order for the positions on your axis (row position 1 should have its label in the first slot in the vector)
vars_x_axis <- c("Euc_Fl", "Euc_Sm","Peat_Fl","Peat_Sm","Pine_Fl","Pine_Sm","PineNeedles_Fl","PineNeedles_Sm","RedOak_Fl","RedOak_Sm")
vars_y_axis <- c("Pneumonia","Influenza","COVID-19","Bronchitis","ILD","Hypersensitivity","COPD","Asthma")
```

## Generate balloon plot
Remember to change ggplot(size=) and geom_text(label=)
```{r}
balloonplot <- ggplot(input_data, aes(x=factor(col), y=factor(row), color=Exposure, size=Index_ExpSc)) +
  scale_fill_brewer(palette="Dark2") +
  geom_point() +    # plot as points
  geom_text(aes(label=formatC(as.numeric(as.character(round(Index_ExpSc, 3))), digits = 3, format = "f"), x=col + 0.25), size=4) +   # display the value next to the "balloons"
  scale_size(range = c(1, 5.5)) + # controls range of dot sizes
  scale_x_discrete(breaks=1:length(vars_x_axis), labels=vars_x_axis, position='top') +   
  scale_y_discrete(breaks=1:length(vars_y_axis), labels=vars_y_axis) +                 
  theme_bw() +
  theme(axis.line = element_blank(),            
        axis.title = element_blank(),          
        panel.border = element_blank(),         
        panel.grid.major.x = element_blank(),   
        panel.grid.minor.x = element_blank(),
        legend.position = "none",
        axis.text = element_text(size = 10, color = "black"),
        axis.ticks = element_blank())
balloonplot
```

## Save balloon plot
Note that the dimensions of the figure you write out will influence how close the rows and columns are to each other. These output dimensions worked well for me for my balloon plot with 4 columns and 5 rows.
NOTE: Export this PDF to TIFF and upload to BioRender for post-processing.
```{r}
pdf(paste0("5_Figures/", list_name, "_BalloonPlot_Dice_ExpScaled_", filters, ".pdf"), height = 4, width = 13)
balloonplot
invisible(dev.off())
```

## Second, scale overlap_matrix within DISEASES
```{r}
# Create function to min-max scale within columns
minmax <- function(x, na.rm = TRUE) {
    return((x-min(x))/(max(x)-min(x)))
}

# Pivot input_data wider where Disease conditions become column names
Index_scale <- pivot_wider(Index_raw, names_from = Disease, values_from = Index_raw)

# Apply function across Disease columns, pivot back to long, and join back to input_data
Index_scale_minmax <- Index_scale %>%
  mutate(across("Asthma":"Hypersensitivity", minmax)) %>% #Remember you made this minmax function above
  pivot_longer(!Exposure, names_to = "Disease", values_to = "Index_DisSc") %>%
  mutate(Coefficient = "Dice")
Index_input <- full_join(Index_input, Index_scale_minmax) 
Index_input <- Index_input %>%
  relocate(Coefficient, .after = Index_DisSc)

# Save as Dice-specific object
Dice_input <- Index_input
```

## Create input dataframe for balloon plot
```{r}
# Assign numeric values to each of your groups to assign the position (x-y coordinates) they will appear on the balloon plot. 
row_df <- data.frame(Exposure = c("Flaming_eucalyptus", "Smoldering_eucalyptus","Flaming_peat","Smoldering_peat","Flaming_pine","Smoldering_pine","Flaming_pine_needles","Smoldering_pine_needles","Flaming_red_oak","Smoldering_red_oak"), row = c(10,9,8,7,6,5,4,3,2,1))
col_df <- data.frame(Disease = c("Asthma","COPD","Hypersensitivity","ILD","Bronchitis","COVID-19","Influenza","Pneumonia"), col = c(1,2,3,4,5,6,7,8))

# Join row and col positions to the data frame 
input_data <- Index_input %>%
  left_join(row_df, by = "Exposure") %>%
  left_join(col_df, by = "Disease")

# Manually create axes labels (if you want the labels to be different - I think if you don't do this, it will just keep current names)
# Note that these should be in numeric order for the positions on your axis (row position 1 should have its label in the first slot in the vector)
vars_x_axis <- c("Asthma","COPD","Hypersensitivity","ILD","Bronchitis","COVID-19","Influenza","Pneumonia")
vars_y_axis <- c("RedOak_Sm","RedOak_Fl","PineNeedles_Sm","PineNeedles_Fl","Pine_Sm","Pine_Fl","Peat_Sm","Peat_Fl","Euc_Sm","Euc_Fl")
```

## Generate balloon plot
Remember to change ggplot(size=) and geom_text(label=)
```{r}
balloonplot <- ggplot(input_data, aes(x=factor(col), y=factor(row), color=Disease, size=Index_DisSc)) +
  scale_fill_brewer(palette="Dark2") +
  geom_point() +    # plot as points
  geom_text(aes(label=formatC(as.numeric(as.character(round(Index_DisSc, 3))), digits = 3, format = "f"), x=col + 0.25), size=4) +   # display the value next to the "balloons"
  scale_size(range = c(1, 5.5)) + # controls range of dot sizes
  scale_x_discrete(breaks=1:length(vars_x_axis), labels=vars_x_axis, position='top') +   
  scale_y_discrete(breaks=1:length(vars_y_axis), labels=vars_y_axis) +                 
  theme_bw() +
  theme(axis.line = element_blank(),            
        axis.title = element_blank(),          
        panel.border = element_blank(),         
        panel.grid.major.x = element_blank(),   
        panel.grid.minor.x = element_blank(),
        legend.position = "none",
        axis.text = element_text(size = 10, color = "black"),
        axis.ticks = element_blank())
balloonplot
```

# Save balloon plot
Note that the dimensions of the figure you write out will influence how close the rows and columns are to each other. These output dimensions worked well for me for my balloon plot with 4 columns and 5 rows.
NOTE: Export this PDF to TIFF and upload to BioRender for post-processing.
```{r}
pdf(paste0("5_Figures/", list_name, "_BalloonPlot_Dice_DisScaled_", filters, ".pdf"), height = 5, width = 11)
balloonplot
invisible(dev.off())
```

# Make master coefficient index df and save
```{r}
# Make master df
Index_input <- Jaccard_input
Index_input <- bind_rows(Index_input, Overlap_input)
Index_input <- bind_rows(Index_input, Dice_input)

# Save onto file generated above
addWorksheet(OUT, "input_values")
writeData(OUT, sheet = "input_values", x = Index_input, rowNames = FALSE)
saveWorkbook(OUT, paste0("4_CountMatrix_Jaccard/", list_name, "_CountMatrix_MultiMeasures_", filters, ".xlsx"), overwrite = TRUE)
```

# ClusterProfiler
See pathfindR vignette here: https://cran.r-project.org/web/packages/pathfindR/vignettes/intro_vignette.html
See clusterProfiler vignette here (VERY IMPORTANT): https://yulab-smu.top/biomedical-knowledge-mining-book/index.html
See clusterProfiler manuscript here: https://www.nature.com/articles/s41596-024-01020-z
Example input data for clusterProfiler here: https://yulab-smu.top/clusterProfiler_protocol/
```{r}
# Clean NCBI human gene symbol codebook by removing columns we don't need and renaming those we do
ncbi_genes <- read.csv("ncbi_human.tsv", sep = '\t')
ncbi_genes <- ncbi_genes[,1:2]
colnames(ncbi_genes) <- c("entrez", "Human_gene")

# Convert Human_gene to NCBI EntrezID and convert to character vector needed for input
Entrez <- TGCs_10cond1dis %>%
  dplyr::select(Human_gene) #Remove extra columns we don't need
Entrez <- left_join(Entrez, ncbi_genes, by = join_by(Human_gene)) #left_join() to only keep TGCs
Entrez <- Entrez[,2] #Remove all columns besides the EntrezIDs
CP_input <- as.character(Entrez$entrez) #Convert to character vector format needed as input

#Load packages we need for Reactome Over-Representation Analysis (ORA) and visualizations
library(ReactomePA)
library(org.Hs.eg.db)
library(enrichplot)
library(ggplot2)

# Run Reactome ORA using enrichPathway(), which is Reactome-specific
Reactome_ORA_results <- enrichPathway(gene=CP_input, pvalueCutoff = 0.05, readable=TRUE)

# Convert Reactome_ORA_results to data frame so we can save below
res1 <- as.data.frame(Reactome_ORA_results)

# Visualize using traditional bar and dot plots
pBar <- barplot(Reactome_ORA_results, showCategory=20)
pBar2 <- barplot(Reactome_ORA_results, showCategory=10)
pBarQscore <- mutate(Reactome_ORA_results, qscore = -log(p.adjust, base=10)) %>% 
    barplot(x="qscore")
pDot <- dotplot(Reactome_ORA_results, showCategory=10) +
  scale_color_gradient(
    low = "red",      # low p.adjust values (significant) → red
    high = "blue",    # high p.adjust values (non-sig) → blue
    trans = "reverse" # flips the legend so red is at the top
  ) +
  labs(color = "Adjusted p-value")
pDot
pDot2 <- dotplot(Reactome_ORA_results, showCategory=20)

# Convert EntrezIDs back to gene symbols for the CNet figures, and visualize using more complicated plots
Reactome_ORA_results_symbols <- setReadable(Reactome_ORA_results, 'org.Hs.eg.db', 'ENTREZID')
# CNet
pCNet <- cnetplot(Reactome_ORA_results_symbols)
# Heatmap
pHeat <- heatplot(Reactome_ORA_results_symbols, showCategory=5)
pHeat2 <- heatplot(Reactome_ORA_results_symbols, showCategory=10)
# Tree plot
Reactome_ORA_results_symbols2 <- pairwise_termsim(Reactome_ORA_results_symbols) #Calculate pair-wise similarities
pTree <- treeplot(Reactome_ORA_results_symbols2, hclust_method = "average")
pTree

# Enrichment map organizes enriched terms into a network with edges connecting overlapping gene sets. In this way, mutually overlapping gene sets are tend to cluster together, making it easy to identify functional module.
pEnrich <- emapplot(Reactome_ORA_results_symbols2)
pEnrich
```

# Clean Reactome_ORA_results dataframe so GeneRatio doesn't output as dates in csv
```{r}
res1 <- res1 %>%
  separate(GeneRatio,
           into = c("GeneRatio_num", "GeneRatio_denom"),
           sep = "/",
           convert = TRUE) %>%   # converts to numeric instead of character
  separate(BgRatio,
           into = c("BgRatio_num", "BgRatio_denom"),
           sep = "/",
           convert = TRUE) %>%   # converts to numeric instead of character
  mutate(GeneRatio = GeneRatio_num / GeneRatio_denom,
         BgRatio = BgRatio_num / BgRatio_denom) %>%
  relocate(GeneRatio, .before = GeneRatio_num) %>%
  relocate(BgRatio, .before = BgRatio_num)
```

# Save Reactome_ORA_results and plots
```{r}
write_csv(res1, file = paste0("9_ClusterProfiler/", date, "_750_ClusterProfilerReactome_RESULTS.csv"))

ggsave(pBar, file = paste0("9_ClusterProfiler/", date, "_750_ClusterProfilerReactome_Bar20.png"), width = 6, height = 12)
ggsave(pBar2, file = paste0("9_ClusterProfiler/", date, "_750_ClusterProfilerReactome_Bar10.png"), width = 6, height = 6)
ggsave(pBarQscore, file = paste0("9_ClusterProfiler/", date, "_750_ClusterProfilerReactome_BarQscore.png"), width = 6, height = 6)
ggsave(pDot, file = paste0("9_ClusterProfiler/", date, "_750_ClusterProfilerReactome_Dot10.png"), width = 6, height = 6)
ggsave(pDot2, file = paste0("9_ClusterProfiler/", date, "_750_ClusterProfilerReactome_Dot20.png"), width = 6, height = 12)
ggsave(pCNet, file = paste0("9_ClusterProfiler/", date, "_750_ClusterProfilerReactome_CNet.png"), width = 6, height = 4)
ggsave(pHeat, file = paste0("9_ClusterProfiler/", date, "_750_ClusterProfilerReactome_Heat5.png"), width = 6, height = 3)
ggsave(pHeat2, file = paste0("9_ClusterProfiler/", date, "_750_ClusterProfilerReactome_Heat10.png"), width = 8, height = 4)
ggsave(pTree, file = paste0("9_ClusterProfiler/", date, "_750_ClusterProfilerReactome_Tree.png"), width = 15, height = 8)
ggsave(pEnrich, file = paste0("9_ClusterProfiler/", date, "_750_ClusterProfilerReactome_EnrichMap.png"), width = 12, height = 8)
```